---
title: "Details on the experimental design"
author: "Juliette Archambeau"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    # code_fold: hide
    toc: true
    toc_depth: 4
    toc_float:
       collapsed: false
    number_sections: true
    highlight: textmate
---

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<style type="text/css">
div.main-container {
  max-width: 2000px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 600px;
}
```

```{r setup, include=FALSE}
options(width = 300)
library(knitr)
library(xtable)
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
knitr::opts_chunk$set(cache=F)
```



# Number of observations in the entire dataset and in each partition

> In all sites

```{r NbobsAllDataset, warning=F,message=F}
data <- readRDS(file="../../data/AllDataPhenoClimSoil.RDS")

# removing missing structure data
data <- data[!(is.na(data$Q1)),]

# removing missing height data
data <- data[!(is.na(data$height)),]

tab <- as.data.frame(table(data$site))
colnames(tab) <-c("Level","All") 
tab$Level <- str_to_title(tab$Level)

data$site_age <- paste0(data$site, " - ", data$age, " months old")
data$site_age <- str_to_sentence(data$site_age)
tab_add <- as.data.frame(table(data$site_age))
colnames(tab_add) <-c("Level","All") 

tab <- bind_rows(tab,tab_add)

tab <- bind_rows(data.frame(Level="All sites",All=dim(data)[[1]] ), tab)
```

> In the P1 partition 

```{r NbobsP1, warning=F,message=F}
data <- readRDS(file="../../data/TrainP1.RDS")

tab_add <- as.data.frame(table(data$site))
colnames(tab_add) <-c("Level","Train1") 
tab_add$Level <- str_to_title(tab_add$Level)

data$site_age <- paste0(data$site, " - ", data$age, " months old")
data$site_age <- str_to_sentence(data$site_age)
tab_add_2 <- as.data.frame(table(data$site_age))
colnames(tab_add_2) <-c("Level","Train1") 

tab_add <- bind_rows(tab_add,tab_add_2)

tab_add <- bind_rows(data.frame(Level="All sites",Train1=dim(data)[[1]] ), tab_add)
tab <- left_join(tab,tab_add)
tab$Train1  <- tab$Train1 %>% replace_na(0)


data <- readRDS(file="../../data/TestP1.RDS")

tab_add <- as.data.frame(table(data$site))
colnames(tab_add) <-c("Level","Test1") 
tab_add$Level <- str_to_title(tab_add$Level)

data$site_age <- paste0(data$site, " - ", data$age, " months old")
data$site_age <- str_to_sentence(data$site_age)
tab_add_2 <- as.data.frame(table(data$site_age))
colnames(tab_add_2) <-c("Level","Test1") 

tab_add <- bind_rows(tab_add,tab_add_2)

tab_add <- bind_rows(data.frame(Level="All sites",Test1=dim(data)[[1]] ), tab_add)
tab <- left_join(tab,tab_add)
tab$Test1  <- tab$Test1 %>% replace_na(0)
```


> In the P2 partition 


```{r Nbobsp2, warning=F,message=F}
data <- readRDS(file="../../data/TrainP2.RDS")

tab_add <- as.data.frame(table(data$site))
colnames(tab_add) <-c("Level","Train2") 
tab_add$Level <- str_to_title(tab_add$Level)

data$site_age <- paste0(data$site, " - ", data$age, " months old")
data$site_age <- str_to_sentence(data$site_age)
tab_add_2 <- as.data.frame(table(data$site_age))
colnames(tab_add_2) <-c("Level","Train2") 

tab_add <- bind_rows(tab_add,tab_add_2)

tab_add <- bind_rows(data.frame(Level="All sites",Train2=dim(data)[[1]] ), tab_add)
tab <- left_join(tab,tab_add)
tab$Train2  <- tab$Train2 %>% replace_na(0)


data <- readRDS(file="../../data/TestP2.RDS")

tab_add <- as.data.frame(table(data$site))
colnames(tab_add) <-c("Level","Test2") 
tab_add$Level <- str_to_title(tab_add$Level)

data$site_age <- paste0(data$site, " - ", data$age, " months old")
data$site_age <- str_to_sentence(data$site_age)
tab_add_2 <- as.data.frame(table(data$site_age))
colnames(tab_add_2) <-c("Level","Test2") 

tab_add <- bind_rows(tab_add,tab_add_2)

tab_add <- bind_rows(data.frame(Level="All sites",Test2=dim(data)[[1]] ), tab_add)
tab <- left_join(tab,tab_add)
tab$Test2  <- tab$Test2 %>% replace_na(0)
```

> In the P3 partition 


```{r Nbobsp3, warning=F,message=F}
data <- readRDS(file="../../data/TrainP3.RDS")

tab_add <- as.data.frame(table(data$site))
colnames(tab_add) <-c("Level","Train3") 
tab_add$Level <- str_to_title(tab_add$Level)

data$site_age <- paste0(data$site, " - ", data$age, " months old")
data$site_age <- str_to_sentence(data$site_age)
tab_add_3 <- as.data.frame(table(data$site_age))
colnames(tab_add_3) <-c("Level","Train3") 

tab_add <- bind_rows(tab_add,tab_add_3)

tab_add <- bind_rows(data.frame(Level="All sites",Train3=dim(data)[[1]] ), tab_add)
tab <- left_join(tab,tab_add)
tab$Train3  <- tab$Train3 %>% replace_na(0)


data <- readRDS(file="../../data/TestP3.RDS")

tab_add <- as.data.frame(table(data$site))
colnames(tab_add) <-c("Level","Test3") 
tab_add$Level <- str_to_title(tab_add$Level)

data$site_age <- paste0(data$site, " - ", data$age, " months old")
data$site_age <- str_to_sentence(data$site_age)
tab_add_3 <- as.data.frame(table(data$site_age))
colnames(tab_add_3) <-c("Level","Test3") 

tab_add <- bind_rows(tab_add,tab_add_3)

tab_add <- bind_rows(data.frame(Level="All sites",Test3=dim(data)[[1]] ), tab_add)
tab <- left_join(tab,tab_add)
tab$Test3  <- tab$Test3 %>% replace_na(0)
```


```{r NbobsJoin, warning=F,message=F}
row.names(tab) <- tab$Level
tab$Level <- NULL
tab <- round(tab,0)
print(xtable(tab, type = "latex",digits=0), file = "../../tables/ExperimentalDesign/ExpDesignNbObs.tex")
kable(tab)
```




# What proportion of trees died in Caceres and Madrid?

```{r ProportionDeadTreesMadridCaceres}
data <- readRDS(file="../../data/AllDataPhenoClimSoil.RDS")
table(data$site)

# removing missing height data
sub <- data[!(is.na(data$height)),]
table(sub$site)
100-344*100/4272 #  91.9 % of trees died in Caceres
100-1067*100/4272 #  75.0 % of the trees died in Madrid

# removing missing structure data
sub<- data[!(is.na(data$Q1)),]
sub <- sub[!(is.na(sub$height)),]
table(sub$site)
100-340*100/4272 #  92.0 % of trees died in Caceres
100-1046*100/4272 #  75.5 % of trees died in Madrid 
```



# How many items in each level?

```{r NbInEachLevel}
data <- readRDS(file="../../data/AllDataPhenoClimSoil.RDS")

# removing missing structure data
data <- data[!(is.na(data$Q1)),]

# removing missing height data
data <- data[!(is.na(data$height)),]

# Number of observations (= measurements)
length((data$tree))

# Number of trees
length(unique(data$tree))

# Number of genotypes
length(unique(data$clon))

# Number of provenances
length(unique(data$prov))
```

```{r LevelsPerProv,message=F,warning=F}
dfclon <- data %>%  dplyr::select(prov,clon) %>% distinct() %>% group_by(prov) %>% dplyr::count()
colnames(dfclon) <- c("Provenances","Number of clones")
dftree <- data %>%  dplyr::select(prov,tree) %>% distinct() %>% group_by(prov) %>% dplyr::count()
colnames(dftree) <- c("Provenances","Number of trees")

dfobs <- data %>%  group_by(prov) %>% tally()
colnames(dfobs) <- c("Provenances","Number of observations")

df <- inner_join(dfclon,dftree)
df <- inner_join(df,dfobs)
kable(df)

print(xtable(df, type = "latex",digits=0), file = "../../tables/ExperimentalDesign/ExpDesignProvenances.tex", include.rownames=FALSE)
```



# Proportion of belonging to each gene pool for each provenance


Matches:

- Northern africa gene pool => NA => Q1

- Corsican gene pool => C => Q2

- Central Spain gene pool => CS => Q3

- French Atlantic gene pool => FA => Q4

- Iberian Atlantic gene pool => IA => Q5

- South-eastern Spain gene pool => SES => Q6



```{r ProportionBelongingGenePools}
data <- readRDS(file="../../data/AllDataPhenoClimSoil.RDS")

# removing missing structure data
data <- data[!(is.na(data$Q1)),]

# removing missing height data
data <- data[!(is.na(data$height)),]

data <- droplevels(data)
ps <- data %>%
  group_by(prov) %>% 
  summarise_at(vars(paste0(rep("Q",6),1:6)), mean)  # %>% mutate_at(2:7,funs(round(., 3))) 

colnames(ps) <- c("Provenance","NA","C","CS","FA","IA","SES")

print(xtable(ps, type = "latex",digits=3), file = "../../tables/ExperimentalDesign/PopStructureByProv.tex", include.rownames=FALSE)

kable(ps)
```



# Percentage of unique trees

```{r PercentageUniqueTrees}
data <- readRDS(file="../../data/AllDataPhenoClimSoil.RDS")

# removing missing structure data
data <- data[!(is.na(data$Q1)),]

# removing missing height data
data <- data[!(is.na(data$height)),]

length(unique(data$tree)) # There are 12,841 trees.

# Duplicates
dup <- data[duplicated(data$tree),]
length(unique(dup$tree)) # There are 11,002 trees with more than one measurements.
tree.dup <- unique(dup$tree)

uni <- data[!(data$tree %in% tree.dup),]
length(unique(uni$tree)) # There are 1,839 trees with one measurement. 

table(uni$site) # Most of these trees are in Madrid. 

1839*100/12841 # About 14% of the trees have only one measurement.
```


`r knitr::opts_chunk$set(eval = F)`

