---
title: "Exploratory analyses"
author: "Juliette Archambeau"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    # code_fold: hide
    toc: true
    toc_depth: 4
    toc_float:
       collapsed: false
    number_sections: true
    highlight: textmate
---

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<style type="text/css">
div.main-container {
  max-width: 2000px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 600px;
}
```

```{r setup, include=FALSE}
options(width = 300)
library(knitr)
library(readr)
library(dplyr)
library(tidyr)
library(ggbiplot)
library(stringr)
library(ggcorrplot)
library(ggpubr)
knitr::opts_chunk$set(cache=F,echo=F)
```


```{r LoadFunctions}
source("../../../scripts/10_functions/vir_lite.R")
ds=0.7
```


```{r LoadData}
data <- readRDS(file = "../../data/AllDataPhenoClimSoil.RDS")

# removing missing structure data
data <- data[!(is.na(data$Q1)),]

# removing missing height data
data <- data[!(is.na(data$height)),]
```

# Correlations between variables


## In the provenances

```{r  SelectDataProv, warning=F,message=F}
# d.prov <- data %>% 
#   dplyr::select(starts_with("Q"),ends_with("prov"),-prov,-ends_with("sub_prov")) %>% 
#   drop_na() %>% 
#   distinct()%>% 
#   scale() 

d.prov.selected <- data %>% 
  dplyr::select(starts_with("Q"),bio1_prov,bio5_prov,bio12_prov,bio14_prov) %>% 
  drop_na() %>% 
  distinct() %>% 
  scale() %>% 
  as_tibble() %>% 
  rename_at(vars(ends_with("_prov")), funs(str_sub(., 1, -6))) %>% 
  dplyr::rename("max.temp"=bio5,"mean.temp"=bio1,"mean.pre"=bio12,"min.pre"=bio14,
                "Northern Africa"=Q1, "Corsica"=Q2, "Central-Spain"=Q3, "French Atlantic"=Q4, "Iberian Atlantic"=Q5, "South-eastern Spain"=Q6)
```

### Correlation plot

```{r  CorrelogramProv, fig.height=8}
corr <- cor(d.prov.selected)

p <- ggcorrplot(corr, hc.order = TRUE, type = "lower",
     outline.col = "white",lab=T,
     legend.title = "Correlation\ncoefficients",
     tl.cex = 16) + 
  theme(legend.text = element_text(size=15),
        legend.title = element_text(size=15),
        legend.position = c(0.2,0.8))
p
ggsave(p,file="../../figs/SuppInfo/ExploringData/CorrplotProv.png",height=10,width=12)
```

### PCA

```{r PcasProv, fig.height=8,fig.width=8}
pca.prov.selected <- prcomp(d.prov.selected)

p <- ggbiplot(pca.prov.selected,varname.size =5,alpha = 0.3,labels.size = 10,
         varname.adjust = 1.5) +  ylim(-2.5, 1.6) +    xlim(-2.3, 2) + 
  theme_bw() +
  theme(plot.title = element_text(size=18),
        axis.title = element_text(size=18),
        axis.text = element_text(size=12))
p
ggsave(p,file="../../figs/SuppInfo/ExploringData/PcaProv.png",height=10,width=12)
```


## In the test sites

```{r SelectDataSite, message=F}
# d.site <- data %>% 
#   dplyr::select(ends_with("1y_site"),ends_with("summer_min_site"),-tmx_min_1y_site,
#                 -tmn_max_1y_site,-pre_min_1y_site,age) %>%
#   drop_na() %>% 
#   distinct()%>% 
#   scale()

d.site <- data %>% 
  dplyr::select("age",
                "pre_summer_min_site",
                "pre_mean_1y_site",
                "tmn_min_1y_site",
                "tmx_max_1y_site",
                "pre_max_1y_site",
                "tmx_mean_1y_site") %>%
  drop_na() %>% 
  distinct()%>% 
  scale() %>% 
  as_tibble() %>% 
  rename_at(vars(ends_with("1y_site")), funs(str_sub(., 1, -9))) %>% 
  rename_at(vars(ends_with("_site")), funs(str_sub(., 1, -6))) %>% 
  dplyr::rename(mean.tmax=tmx_mean, 
         max.tmax=tmx_max, 
         min.presummer=pre_summer_min,
         min.tmn=tmn_min,
         mean.pre=pre_mean,
         max.pre=pre_max)
```

### Correlation plot

```{r CorrSite, fig.height=8}
cor <- cor(d.site)

p <- ggcorrplot(cor, 
                hc.order = TRUE, 
                type = "lower",
     outline.col = "white",
     lab=T,
     legend.title = "Correlation\ncoefficients",
     tl.cex = 16
     ) + 
  theme(legend.text = element_text(size=15),
        legend.title = element_text(size=15),
        legend.position = c(0.2,0.8))
p
ggsave(p,file="../../figs/SuppInfo/ExploringData/CorrplotSite.png",height=10,width=12)
```

### PCA

```{r PcasSite, fig.height=9, fig.width=8}
pca.site <- prcomp(d.site)

p <- ggbiplot(pca.site,varname.size =5,labels.size = 10,
         varname.adjust = 1.5) +   ylim(-2, 2) +    xlim(-3, 3) + 
  theme_bw() + geom_point(size=3,alpha = 0.5) +
  theme(plot.title = element_text(size=18),
        axis.title = element_text(size=18),
        axis.text = element_text(size=12))
p
ggsave(p,file="../../figs/SuppInfo/ExploringData/PcaSite.png",height=10,width=12)
```


# Distribution of heights in Caceres and Madrid

> No strong footprints of selection on height variations in Caceres and Madrid


```{r DistributionHeightMadridCaceres,message=F, fig.width=10}
p <- ggarrange(
data %>% dplyr::filter(site=="caceres") %>% 
  ggplot() + geom_histogram(aes(x=height),color="grey80", fill=vir_lite("pink",ds=ds),bins=50) + theme_bw() +
  labs(x="Height (mm)",y="Count"),

data %>% dplyr::filter(site=="madrid") %>% 
  ggplot() + geom_histogram(aes(x=height),color="grey80", fill=vir_lite("deeppink",ds=ds),bins=50) + theme_bw() +
  labs(x="Height (mm)",y=""),

nrow=1,labels=c("A) Caceres","B) Madrid"),font.label = list(size=20))

p
ggsave(p, file="../../figs/SuppInfo/ExploringData/DistributionHeightCaceresMadrid.png",width=14)
```

                             
`r knitr::opts_chunk$set(eval = F)`

