---
title: "Predictive Models - Figures"
author: "Juliette Archambeau"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    # code_fold: hide
    toc: true
    toc_depth: 5
    toc_float:
       collapsed: false
    number_sections: true
    theme: paper
    highlight: textmate
---


<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<style type="text/css">
body{ /* Normal  */
      font-size: 16px;
  }
div.main-container {
  max-width: 2000px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 600px;
}
```

```{r setup, include=FALSE}
library(knitr)
options(width = 300)
knitr::opts_chunk$set(fig.width = 8, fig.height = 5, cache = F,echo=F)
library(rstan)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
library(bayesplot)
library(gridExtra)
library(grid)
library(ggplot2)
library(ggridges)
library(latex2exp)
color_scheme_set("green")
library(tibble)
library(ggpubr)
library(cowplot)
library(rethinking)
library(brms)
library(dplyr)
library(stringr)
library(tidyr)
```

<!-- Load models -->

```{r Partition}
# Data partition used to fit the models:
part <- "P1"
```

<!-- Functions used -->

```{r FuntionsUsed,echo=F}
source("../../scripts/Functions/vir_lite.R")

# Paramters
ds=0.7
```


# Posterior distribution of M7 and M8


```{r LoadModels}
models <- list()
models[[1]] <- readRDS(file=paste0("../../outputs/models/",part,"/MOD7.rds"))
models[[2]] <- readRDS(file=paste0("../../outputs/models/",part,"/MOD8.rds"))
names(models) <- c("M7","M8")
```


```{r ListFigs}
figs.GP <- list()
figs.beta <- list()
```

## Population structure 

```{r figsGP, fig.width=20,fig.height=8,message=F,warning=F}
for (i in 1:2){
POST <- posterior_samples(models[[i]],pars = "^r_mmQ1Q2Q3Q4Q5Q6\\[")
colnames(POST) <- str_sub(colnames(POST),18,-12)
POST <- as.data.frame(t(POST))
POST$genepool <- as.factor(rownames(POST))

posteriorsimpelmodellong <- POST %>%  as_tibble() %>% 
gather(key = "key", value = "value", -genepool)%>%
  group_by(genepool) %>%
  dplyr::mutate(meanpergenepool = mean(value))%>%
  ungroup()

figs.GP[[i]] <- ggplot()+
  geom_vline(xintercept = 0, 
             col        = "grey70") +
  stat_density_ridges(data  = posteriorsimpelmodellong, 
                                aes(x      = value,
                                    y      = reorder(as.factor(genepool), meanpergenepool),
                                    fill   = as.factor(genepool),
                                    vline_color = ..quantile..),
                                scale = 2, 
                                alpha = .6,
                                rel_min_height=c(.0044),
                                size=0.5,
                                quantile_lines = TRUE, quantiles = c(0.025,0.5,0.975)) +
        scale_discrete_manual("vline_color",
                        values = c("blue", "red", "blue", "black"), 
                        breaks = c(1, 2),
                        labels = c("2.5 and 97.5th percentiles", "Median"),
                        name = NULL) +
  scale_y_discrete(labels=c("Q1"=parse(text = TeX("$g_{NA}$")),
                            "Q2"=parse(text = TeX("$g_{C}$")),
                            "Q3"=parse(text = TeX("$g_{CS}$")),
                            "Q4"=parse(text = TeX("$g_{FA}$")),
                            "Q5"=parse(text = TeX("$g_{IA}$")),
                            "Q6"=parse(text = TeX("$g_{SES}$")))) +
  coord_cartesian(c(-0.5,0.6))+
  scale_fill_manual(values=c("orangered3",
                             "gold2",
                             "darkorchid3",
                             "navyblue",
                             "turquoise2",
                             "green3"), labels = c("Northern Africa (NA)", 
                                                   "Corsica (C)",
                                                   "Central Spain (CS)",
                                                   "French Atlantic (FA)",
                                                   "Iberian Atlantic (IA)",
                                                   "South-eastern Spain (SES)")) +
  labs(fill     = "Gene pools",
       y        = "", 
       x = ""
       #x        = TeX("Estimate of the gene pool intercept g_{j}")
       ) + 
  theme_bw() + theme(axis.text = element_text(size=22),axis.title = element_text(size=22), 
                        legend.text = element_text(size=16),legend.title = element_text(size=17))
}

figs.GP[[1]] <- figs.GP[[1]] + theme(legend.position = "none")
pGP <- ggarrange(figs.GP[[1]],figs.GP[[2]],labels=c("M7. a)","M8. a)"),font.label = list(size = 20),hjust=c(-0.1,-0.1),vjust=c(1.6,1.6),nrow=1,widths = c(1,1.3))
```

## Provenance climates and PEAs

```{r figsbeta, fig.width=20,fig.height=8,message=F,warning=F}
# site <- unique(data$site)
# 
# par <-paste(paste0("'r_site[",site,",",variable,"]'")) 
# par2 <-paste0("beta_PEA_",str_to_title(site))
# tocopy <- noquote(str_sub(stri_paste(paste0(par2," = ",par,","),collapse = ""),0,-2))
# 
# par <-paste(paste0("'r_site[",site,",bio14_prov.sc]'")) 
# par2 <-paste0("beta_MinPre_",str_to_title(site))
# tocopy <- noquote(str_sub(stri_paste(paste0(par2," = ",par,","),collapse = ""),0,-2))
# 
# par <-paste(paste0("'r_site[",site,",bio5_prov.sc]'")) 
# par2 <-paste0("beta_MaxTemp_",str_to_title(site))
# tocopy <- noquote(str_sub(stri_paste(paste0(par2," = ",par,","),collapse = ""),0,-2))

for(i in 1:2){
  
if(i==1){
  pea="gPEA"
  variable ="gPEA.sc"
} else if (i==2){
  pea="rPEA"
  variable ="rPEA.sc"
}
  
  
POST <- posterior_samples(models[[i]],pars = "^r_site\\[.*sc\\]") %>% dplyr::rename(
  beta_PEA_Portugal = paste0('r_site[portugal,',variable,']'),
  beta_PEA_Bordeaux = paste0('r_site[bordeaux,',variable,']'),
  beta_PEA_Asturias = paste0('r_site[asturias,',variable,']'),
  beta_PEA_Madrid = paste0('r_site[madrid,',variable,']'),
  beta_PEA_Caceres = paste0('r_site[caceres,',variable,']'),
  beta_MinPre_Portugal = 'r_site[portugal,bio14_prov.sc]',
  beta_MinPre_Bordeaux = 'r_site[bordeaux,bio14_prov.sc]',
  beta_MinPre_Asturias = 'r_site[asturias,bio14_prov.sc]',
  beta_MinPre_Madrid = 'r_site[madrid,bio14_prov.sc]',
  beta_MinPre_Caceres = 'r_site[caceres,bio14_prov.sc]',
  beta_MaxTemp_Portugal = 'r_site[portugal,bio5_prov.sc]',
  beta_MaxTemp_Bordeaux = 'r_site[bordeaux,bio5_prov.sc]',
  beta_MaxTemp_Asturias = 'r_site[asturias,bio5_prov.sc]',
  beta_MaxTemp_Madrid = 'r_site[madrid,bio5_prov.sc]',
  beta_MaxTemp_Caceres = 'r_site[caceres,bio5_prov.sc]'
  )
POST <- as.data.frame(t(POST))
POST$var <- as.factor(rownames(POST))

posteriorsimpelmodellong <- POST %>%  as_tibble() %>% 
gather(key = "key", value = "value", -var)


figs.beta[[i]] <- ggplot()+
  geom_vline(xintercept = 0, col="grey70") +
  stat_density_ridges(data  = posteriorsimpelmodellong, 
                                aes(x      = value,
                                    y      = var,
                                    fill   = as.factor(var),
                                    vline_color = ..quantile..),
                                scale = 1.8, 
                                alpha = .8,
                                size=0.5,
                                rel_min_height=.01,
                                quantile_lines = TRUE, 
                                quantiles = c(0.025,0.5,0.975)) +
  # coord_cartesian(c(-,0.8))+
      scale_discrete_manual("vline_color",
                        values = c("blue", "red", "blue", "black"), 
                        breaks = c(1, 2),
                        labels = c("5% & 95% quantiles", "mean"),
                        name = NULL) +
  scale_y_discrete(labels=c("beta_MaxTemp_Caceres"=parse(text = TeX("$\\beta_{max.temp,Caceres}$")),
                            "beta_MaxTemp_Bordeaux"=parse(text = TeX("$\\beta_{max.temp,Bordeaux}$")),
                            "beta_MaxTemp_Portugal"=parse(text = TeX("$\\beta_{max.temp,Portugal}$")),
                            "beta_MaxTemp_Madrid"=parse(text = TeX("$\\beta_{max.temp,Madrid}$")),
                            "beta_MaxTemp_Asturias"=parse(text = TeX("$\\beta_{max.temp,Asturias}$")),
                            "beta_MinPre_Caceres"=parse(text = TeX("$\\beta_{min.pre,Caceres}$")),
                            "beta_MinPre_Bordeaux"=parse(text = TeX("$\\beta_{min.pre,Bordeaux}$")),
                            "beta_MinPre_Portugal"=parse(text = TeX("$\\beta_{min.pre,Portugal}$")),
                            "beta_MinPre_Madrid"=parse(text = TeX("$\\beta_{min.pre,Madrid}$")),
                            "beta_MinPre_Asturias"=parse(text = TeX("$\\beta_{min.pre,Asturias}$")),
                            "beta_PEA_Caceres"=parse(text = TeX(paste0("$\\beta_{",pea,",Caceres}$"))),
                            "beta_PEA_Madrid"=parse(text = TeX(paste0("$\\beta_{",pea,",Madrid}$"))),
                            "beta_PEA_Portugal"=parse(text = TeX(paste0("$\\beta_{",pea,",Portugal}$"))),
                            "beta_PEA_Asturias"=parse(text = TeX(paste0("$\\beta_{",pea,",Asturias}$"))),
                            "beta_PEA_Bordeaux"=parse(text = TeX(paste0("$\\beta_{",pea,",Bordeaux}$")))
                            )) + 
  labs(y = "", 
       x = ""
       #x        = TeX(paste0("Estimate of the varying slopes $\\beta_{",pea,",site}$, $\\beta_{min.pre,site}$ and $\\beta_{max.temp,site}$"))
       ) + 
  scale_fill_manual(values=c(vir_lite("cyan2",ds=ds),
                             vir_lite("navyblue",ds=ds),
                             vir_lite("pink",ds=ds),
                             vir_lite("deeppink",ds=ds),
                             vir_lite("dodgerblue2",ds=ds),
                             vir_lite("cyan2",ds=ds),
                             vir_lite("navyblue",ds=ds),
                             vir_lite("pink",ds=ds),
                             vir_lite("deeppink",ds=ds),
                             vir_lite("dodgerblue2",ds=ds),
                             vir_lite("cyan2",ds=ds),
                             vir_lite("navyblue",ds=ds),
                             vir_lite("pink",ds=ds),
                             vir_lite("deeppink",ds=ds),
                             vir_lite("dodgerblue2",ds=ds))) +
  theme_bw() + theme(axis.text = element_text(size=22),axis.title = element_text(size=22),
                     legend.position = "none",
                    legend.text = element_text(size=16),legend.title = element_text(size=17))
  }

pbeta <- ggarrange(figs.beta[[1]],figs.beta[[2]],labels=c("M7. b)","M8. b)"),font.label = list(size = 20),hjust=c(-0.1,-0.1),vjust=c(1.6,1.6),nrow=1)
```

## Figures in the supplementary information 

```{r figtot, fig.width=20,fig.height=14,message=F,warning=F}
figtot <- ggarrange(pGP,pbeta,nrow=2,heights=c(1,2)) 
figtot
ggsave(figtot,file=paste0("../../figs/SuppInfo/M7M8Posteriors",part,".png"),height=12,width=20)
```

<!-- ## Figure in the manuscript (finally not used in the last version of the manuscript) -->

<!-- ```{r figM8, fig.width=10,fig.height=14,message=F,warning=F} -->
<!-- if(part == "P1"){ -->
<!-- figM8 <- ggarrange(figs.GP[[2]], figs.beta[[2]],labels=c("A)","B)"), -->
<!--           font.label = list(size = 20),hjust=c(-0.1,-0.1),vjust=c(1.6,1.6),nrow=2) -->
<!-- ggsave(figM8,file=paste0("../../figs/manuscript/M8Posteriors",part,".png"),height=14,width=11)   -->
<!-- } -->
<!-- ``` -->

##  Figure 1 bis (not used in the manuscript) 

Done the 01/06/2020.

The posterior distributions of the population structure and the site-specific slopes of the climatic variables and PEAs are too different to be ploted together.


```{r FigsAll, fig.width=20,fig.height=15,message=F,warning=F}

figs.all <- list()

for(i in 1:2){
  
if(i==1){
  pea="gPEA"
  variable ="gPEA.sc"
} else if (i==2){
  pea="rPEA"
  variable ="rPEA.sc"
}
  
  
POST <- posterior_samples(models[[i]],pars = "^(r_site\\[.*sc\\]|r_mmQ1Q2Q3Q4Q5Q6\\[)") %>% dplyr::rename(
  beta_PEA_Portugal = paste0('r_site[portugal,',variable,']'),
  beta_PEA_Bordeaux = paste0('r_site[bordeaux,',variable,']'),
  beta_PEA_Asturias = paste0('r_site[asturias,',variable,']'),
  beta_PEA_Madrid = paste0('r_site[madrid,',variable,']'),
  beta_PEA_Caceres = paste0('r_site[caceres,',variable,']'),
  beta_MinPre_Portugal = 'r_site[portugal,bio14_prov.sc]',
  beta_MinPre_Bordeaux = 'r_site[bordeaux,bio14_prov.sc]',
  beta_MinPre_Asturias = 'r_site[asturias,bio14_prov.sc]',
  beta_MinPre_Madrid = 'r_site[madrid,bio14_prov.sc]',
  beta_MinPre_Caceres = 'r_site[caceres,bio14_prov.sc]',
  beta_MaxTemp_Portugal = 'r_site[portugal,bio5_prov.sc]',
  beta_MaxTemp_Bordeaux = 'r_site[bordeaux,bio5_prov.sc]',
  beta_MaxTemp_Asturias = 'r_site[asturias,bio5_prov.sc]',
  beta_MaxTemp_Madrid = 'r_site[madrid,bio5_prov.sc]',
  beta_MaxTemp_Caceres = 'r_site[caceres,bio5_prov.sc]',
  g_NA = "r_mmQ1Q2Q3Q4Q5Q6[Q1,Intercept]",
  g_C = "r_mmQ1Q2Q3Q4Q5Q6[Q2,Intercept]",
  g_CS = "r_mmQ1Q2Q3Q4Q5Q6[Q3,Intercept]",
  g_FA = "r_mmQ1Q2Q3Q4Q5Q6[Q4,Intercept]",
  g_IA = "r_mmQ1Q2Q3Q4Q5Q6[Q5,Intercept]",
  g_SES = "r_mmQ1Q2Q3Q4Q5Q6[Q6,Intercept]"
  )
POST <- as.data.frame(t(POST))
POST$var <- as.factor(rownames(POST))

posteriorsimpelmodellong <- POST %>%  as_tibble() %>% 
gather(key = "key", value = "value", -var)


figs.all[[i]] <- ggplot()+
  geom_vline(xintercept = 0, col="grey70") +
  stat_density_ridges(data  = posteriorsimpelmodellong, 
                                aes(x      = value,
                                    y      = var,
                                    fill   = as.factor(var),
                                    vline_color = ..quantile..),
                                scale = 1.8, 
                                alpha = .8,
                                size=0.8,
                                rel_min_height=.001,
                                quantile_lines = TRUE, 
                                quantiles = c(0.025,0.5,0.975)) +
  coord_cartesian(c(-0.35,0.45))+
      scale_discrete_manual("vline_color",
                        values = c("blue", "red", "blue", "black"), 
                        breaks = c(1, 2),
                        labels = c("5% & 95% quantiles", "mean"),
                        name = NULL) +
  scale_y_discrete(labels=c("beta_MaxTemp_Caceres"=parse(text = TeX("$\\beta_{max.temp,Caceres}$")),
                            "beta_MaxTemp_Bordeaux"=parse(text = TeX("$\\beta_{max.temp,Bordeaux}$")),
                            "beta_MaxTemp_Portugal"=parse(text = TeX("$\\beta_{max.temp,Portugal}$")),
                            "beta_MaxTemp_Madrid"=parse(text = TeX("$\\beta_{max.temp,Madrid}$")),
                            "beta_MaxTemp_Asturias"=parse(text = TeX("$\\beta_{max.temp,Asturias}$")),
                            "beta_MinPre_Caceres"=parse(text = TeX("$\\beta_{min.pre,Caceres}$")),
                            "beta_MinPre_Bordeaux"=parse(text = TeX("$\\beta_{min.pre,Bordeaux}$")),
                            "beta_MinPre_Portugal"=parse(text = TeX("$\\beta_{min.pre,Portugal}$")),
                            "beta_MinPre_Madrid"=parse(text = TeX("$\\beta_{min.pre,Madrid}$")),
                            "beta_MinPre_Asturias"=parse(text = TeX("$\\beta_{min.pre,Asturias}$")),
                            "beta_PEA_Caceres"=parse(text = TeX(paste0("$\\beta_{",pea,",Caceres}$"))),
                            "beta_PEA_Madrid"=parse(text = TeX(paste0("$\\beta_{",pea,",Madrid}$"))),
                            "beta_PEA_Portugal"=parse(text = TeX(paste0("$\\beta_{",pea,",Portugal}$"))),
                            "beta_PEA_Asturias"=parse(text = TeX(paste0("$\\beta_{",pea,",Asturias}$"))),
                            "beta_PEA_Bordeaux"=parse(text = TeX(paste0("$\\beta_{",pea,",Bordeaux}$"))),
                            "g_NA" = parse(text = TeX(paste0("$g_{NA}$"))),
                            "g_C" = parse(text = TeX(paste0("$g_{C}$"))),
                            "g_CS" =parse(text = TeX(paste0("$g_{CS}$"))),
                            "g_FA" = parse(text = TeX(paste0("$g_{FA}$"))),
                            "g_IA" = parse(text = TeX(paste0("$g_{IA}$"))),
                            "g_SES" = parse(text = TeX(paste0("$g_{SES}$")))
                            )) + 
  labs(title="",
       y        = "", 
       x = "Parameter estimates"
       #x        = TeX(paste0("Estimate of the varying slopes $\\beta_{",pea,",site}$, $\\beta_{min.pre,site}$ and $\\beta_{max.temp,site}$"))
       ) + 
  scale_fill_manual(values=c(vir_lite("cyan2",ds=ds),
                             vir_lite("navyblue",ds=ds),
                             vir_lite("pink",ds=ds),
                             vir_lite("deeppink",ds=ds),
                             vir_lite("dodgerblue2",ds=ds),
                             vir_lite("cyan2",ds=ds),
                             vir_lite("navyblue",ds=ds),
                             vir_lite("pink",ds=ds),
                             vir_lite("deeppink",ds=ds),
                             vir_lite("dodgerblue2",ds=ds),
                             vir_lite("cyan2",ds=ds),
                             vir_lite("navyblue",ds=ds),
                             vir_lite("pink",ds=ds),
                             vir_lite("deeppink",ds=ds),
                             vir_lite("dodgerblue2",ds=ds),
                             "orangered3",
                             "gold2",
                             "darkorchid3",
                             "navyblue",
                             "turquoise2",
                             "green3"
                             )) +
  theme_bw() + theme(axis.text = element_text(size=22),axis.title = element_text(size=22), 
                    legend.position = "none", plot.title = element_text(size=22)) 
  }
figs.all[[1]]
figs.all[[2]]
```



# Posterior distributions in M9 to M12 (in the manuscript)

## M9 => Pop structure

```{r loadM9}
mod <- readRDS(file=paste0("../../outputs/models/",part,"/MOD9.rds"))
```

```{r M9PostGenePoolIntercepts, fig.width=8,warning=F,message=F}
POST <- posterior_samples(mod,pars = "^r_mmQ1Q2Q3Q4Q5Q6\\[")
colnames(POST) <- str_sub(colnames(POST),18,-12)
POST <- as.data.frame(t(POST))
POST$genepool <- as.factor(rownames(POST))

posteriorsimpelmodellong <- POST %>%  as_tibble() %>% 
gather(key = "key", value = "value", -genepool)%>%
  group_by(genepool) %>%
  dplyr::mutate(meanpergenepool = mean(value))%>%
  ungroup()

pGP <- ggplot()+
  geom_vline(xintercept = 0, 
             col        = "grey70") +
  stat_density_ridges(data  = posteriorsimpelmodellong, 
                                aes(x      = value,
                                    y      = reorder(as.factor(genepool), meanpergenepool),
                                    fill   = as.factor(genepool),
                                    vline_color = ..quantile..),
                                scale = 2, 
                                alpha = .6,
                                rel_min_height=c(.001),
                                size=0.5,
                                quantile_lines = TRUE, quantiles = c(0.025,0.5,0.975)) +
        scale_discrete_manual("vline_color",
                        values = c("blue", "red", "blue", "black"), 
                        breaks = c(1, 2),
                        labels = c("2.5 and 97.5th percentiles", "Median"),
                        name = NULL) +
  coord_cartesian(c(-0.5,0.4))+
  scale_fill_manual(values=c("orangered3",
                             "gold2",
                             "darkorchid3",
                             "navyblue",
                             "turquoise2",
                             "green3"), labels = c("Northern Africa (NA)", 
                                                   "Corsica (C)",
                                                   "Central Spain (CS)",
                                                   "French Atlantic (FA)",
                                                   "Iberian Atlantic (IA)",
                                                   "South-eastern Spain (SES)")) +
  scale_y_discrete(labels=c("Q1"=parse(text = TeX("$g_{NA}$")),
                            "Q2"=parse(text = TeX("$g_{C}$")),
                            "Q3"=parse(text = TeX("$g_{CS}$")),
                            "Q4"=parse(text = TeX("$g_{FA}$")),
                            "Q5"=parse(text = TeX("$g_{IA}$")),
                            "Q6"=parse(text = TeX("$g_{SES}$")))) + 
  labs(fill     = "Gene pools",
       y        = "", 
       x=""#x        = TeX("Gene pool intercept g_{j}")
       ) + 
  theme_bw() + theme(axis.text = element_text(size=18),axis.title = element_text(size=18), 
                        legend.text = element_text(size=16),legend.title = element_text(size=17))
```

## M10 => Climate in the provenances

```{r loadM10}
mod <- readRDS(file=paste0("../../outputs/models/",part,"/MOD10.rds"))
```

```{r FigM10, fig.width=20,fig.height=8,message=F,warning=F}
POST <- posterior_samples(mod,pars = "^r_site\\[.*sc\\]") %>% dplyr::rename(
  beta_MinPre_Portugal = 'r_site[portugal,bio14_prov.sc]',
  beta_MinPre_Bordeaux = 'r_site[bordeaux,bio14_prov.sc]',
  beta_MinPre_Asturias = 'r_site[asturias,bio14_prov.sc]',
  beta_MinPre_Madrid = 'r_site[madrid,bio14_prov.sc]',
  beta_MinPre_Caceres = 'r_site[caceres,bio14_prov.sc]',
  beta_MaxTemp_Portugal = 'r_site[portugal,bio5_prov.sc]',
  beta_MaxTemp_Bordeaux = 'r_site[bordeaux,bio5_prov.sc]',
  beta_MaxTemp_Asturias = 'r_site[asturias,bio5_prov.sc]',
  beta_MaxTemp_Madrid = 'r_site[madrid,bio5_prov.sc]',
  beta_MaxTemp_Caceres = 'r_site[caceres,bio5_prov.sc]'
  )
POST <- as.data.frame(t(POST))
POST$var <- as.factor(rownames(POST))

posteriorsimpelmodellong <- POST %>%  as_tibble() %>% 
gather(key = "key", value = "value", -var)


pCP <- ggplot()+
  geom_vline(xintercept = 0, col="grey70") +
  stat_density_ridges(data  = posteriorsimpelmodellong, 
                                aes(x      = value,
                                    y      = var,
                                    fill   = as.factor(var),
                                    vline_color = ..quantile..),
                                scale = 1.8, 
                                alpha = .8,
                                size=0.5,
                                rel_min_height=.01,
                                quantile_lines = TRUE, 
                                quantiles = c(0.025,0.5,0.975)) +
  # coord_cartesian(c(-,0.8))+
      scale_discrete_manual("vline_color",
                        values = c("blue", "red", "blue", "black"), 
                        breaks = c(1, 2),
                        labels = c("5% & 95% quantiles", "mean"),
                        name = NULL) +
  scale_y_discrete(labels=c("beta_MaxTemp_Caceres"=parse(text = TeX("$\\beta_{max.temp,Caceres}$")),
                            "beta_MaxTemp_Bordeaux"=parse(text = TeX("$\\beta_{max.temp,Bordeaux}$")),
                            "beta_MaxTemp_Portugal"=parse(text = TeX("$\\beta_{max.temp,Portugal}$")),
                            "beta_MaxTemp_Madrid"=parse(text = TeX("$\\beta_{max.temp,Madrid}$")),
                            "beta_MaxTemp_Asturias"=parse(text = TeX("$\\beta_{max.temp,Asturias}$")),
                            "beta_MinPre_Caceres"=parse(text = TeX("$\\beta_{min.pre,Caceres}$")),
                            "beta_MinPre_Bordeaux"=parse(text = TeX("$\\beta_{min.pre,Bordeaux}$")),
                            "beta_MinPre_Portugal"=parse(text = TeX("$\\beta_{min.pre,Portugal}$")),
                            "beta_MinPre_Madrid"=parse(text = TeX("$\\beta_{min.pre,Madrid}$")),
                            "beta_MinPre_Asturias"=parse(text = TeX("$\\beta_{min.pre,Asturias}$")))) + 
  labs(y        = "", 
       x="") + #x = "Parameter estimates" ) + 
  scale_fill_manual(values=c(vir_lite("cyan2",ds=ds),
                             vir_lite("navyblue",ds=ds),
                             vir_lite("pink",ds=ds),
                             vir_lite("deeppink",ds=ds),
                             vir_lite("dodgerblue2",ds=ds),
                             vir_lite("cyan2",ds=ds),
                             vir_lite("navyblue",ds=ds),
                             vir_lite("pink",ds=ds),
                             vir_lite("deeppink",ds=ds),
                             vir_lite("dodgerblue2",ds=ds))) +
  theme_bw() + theme(axis.text = element_text(size=22),axis.title = element_text(size=22), 
                    legend.position = "none", plot.title = element_text(size=22)) 
```

## M11 => gPEAs

```{r loadM11}
mod <- readRDS(file=paste0("../../outputs/models/",part,"/MOD11.rds"))
```

```{r FigsM11, fig.width=20,fig.height=6,message=F,warning=F}
pea="gPEA"
variable ="gPEA.sc"
  
POST <- posterior_samples(mod,pars = "^r_site\\[.*sc\\]") %>% dplyr::rename(
  beta_PEA_Portugal = paste0('r_site[portugal,',variable,']'),
  beta_PEA_Bordeaux = paste0('r_site[bordeaux,',variable,']'),
  beta_PEA_Asturias = paste0('r_site[asturias,',variable,']'),
  beta_PEA_Madrid = paste0('r_site[madrid,',variable,']'),
  beta_PEA_Caceres = paste0('r_site[caceres,',variable,']')
  )
POST <- as.data.frame(t(POST))
POST$var <- as.factor(rownames(POST))

posteriorsimpelmodellong <- POST %>%  as_tibble() %>% 
gather(key = "key", value = "value", -var)


pgpea <- ggplot()+
  geom_vline(xintercept = 0, col="grey70") +
  stat_density_ridges(data  = posteriorsimpelmodellong, 
                                aes(x      = value,
                                    y      = var,
                                    fill   = as.factor(var),
                                    vline_color = ..quantile..),
                                scale = 1.8, 
                                alpha = .8,
                                size=0.5,
                                rel_min_height=.01,
                                quantile_lines = TRUE, 
                                quantiles = c(0.025,0.5,0.975)) +
  # coord_cartesian(c(-,0.8))+
      scale_discrete_manual("vline_color",
                        values = c("blue", "red", "blue", "black"), 
                        breaks = c(1, 2),
                        labels = c("5% & 95% quantiles", "mean"),
                        name = NULL) +
  scale_y_discrete(labels=c("beta_PEA_Caceres"=parse(text = TeX(paste0("$\\beta_{",pea,",Caceres}$"))),
                            "beta_PEA_Madrid"=parse(text = TeX(paste0("$\\beta_{",pea,",Madrid}$"))),
                            "beta_PEA_Portugal"=parse(text = TeX(paste0("$\\beta_{",pea,",Portugal}$"))),
                            "beta_PEA_Asturias"=parse(text = TeX(paste0("$\\beta_{",pea,",Asturias}$"))),
                            "beta_PEA_Bordeaux"=parse(text = TeX(paste0("$\\beta_{",pea,",Bordeaux}$")))
                            )) + 
  labs(fill = "Sites",
       y = "", 
       x="") + #x = TeX(paste0("Estimate of parameters $\\beta_{",pea,",s}$"))) + 
  scale_fill_manual(values=c(vir_lite("cyan2",ds=ds),
                             vir_lite("navyblue",ds=ds),
                             vir_lite("pink",ds=ds),
                             vir_lite("deeppink",ds=ds),
                             vir_lite("dodgerblue2",ds=ds)),
                    labels = c("Asturias", 
                               "Bordeaux",
                               "Caceres",
                               "Madrid",
                               "Portugal")) +
  theme_bw() + theme(axis.text = element_text(size=22),axis.title = element_text(size=22), 
                    legend.text = element_text(size=16),legend.title = element_text(size=17),
                    #legend.position = "none", 
                    plot.title = element_text(size=22))   + guides(vline_color = FALSE)
```

## M12 => rPEAs

```{r loadM12}
mod <- readRDS(file=paste0("../../outputs/models/",part,"/MOD12.rds"))
```

```{r figsbetaM12, fig.width=20,fig.height=6,message=F,warning=F}
pea="rPEA"
variable ="rPEA.sc"
  
POST <- posterior_samples(mod,pars = "^r_site\\[.*sc\\]") %>% dplyr::rename(
  beta_PEA_Portugal = paste0('r_site[portugal,',variable,']'),
  beta_PEA_Bordeaux = paste0('r_site[bordeaux,',variable,']'),
  beta_PEA_Asturias = paste0('r_site[asturias,',variable,']'), 
  beta_PEA_Madrid = paste0('r_site[madrid,',variable,']'),
  beta_PEA_Caceres = paste0('r_site[caceres,',variable,']')
  )
POST <- as.data.frame(t(POST))
POST$var <- as.factor(rownames(POST))

posteriorsimpelmodellong <- POST %>%  as_tibble() %>% 
gather(key = "key", value = "value", -var)


prpea <- ggplot()+
  geom_vline(xintercept = 0, col="grey70") +
  stat_density_ridges(data  = posteriorsimpelmodellong, 
                                aes(x      = value,
                                    y      = var,
                                    fill   = as.factor(var),
                                    vline_color = ..quantile..),
                                scale = 1.8, 
                                alpha = .8,
                                size=0.5,
                                rel_min_height=.01,
                                quantile_lines = TRUE, 
                                quantiles = c(0.025,0.5,0.975)) +
  # coord_cartesian(c(-,0.8))+
      scale_discrete_manual("vline_color",
                        values = c("blue", "red", "blue", "black"), 
                        breaks = c(1, 2),
                        labels = c("5% & 95% quantiles", "mean"),
                        name = NULL) +
  scale_y_discrete(labels=c("beta_PEA_Caceres"=parse(text = TeX(paste0("$\\beta_{",pea,",Caceres}$"))),
                            "beta_PEA_Madrid"=parse(text = TeX(paste0("$\\beta_{",pea,",Madrid}$"))),
                            "beta_PEA_Portugal"=parse(text = TeX(paste0("$\\beta_{",pea,",Portugal}$"))),
                            "beta_PEA_Asturias"=parse(text = TeX(paste0("$\\beta_{",pea,",Asturias}$"))),
                            "beta_PEA_Bordeaux"=parse(text = TeX(paste0("$\\beta_{",pea,",Bordeaux}$")))
                            )) + 
  labs(title="",
       y        = "", 
       x=""#x = TeX(paste0("Estimate of parameters $\\beta_{",pea,",s}$"))
       ) + 
  scale_fill_manual(values=c(vir_lite("cyan2",ds=ds),
                             vir_lite("navyblue",ds=ds),
                             vir_lite("pink",ds=ds),
                             vir_lite("deeppink",ds=ds),
                             vir_lite("dodgerblue2",ds=ds))) +
  theme_bw() + theme(axis.text = element_text(size=22),axis.title = element_text(size=22), 
                    legend.position = "none", plot.title = element_text(size=22)) 
```

```{r FigFinal2,warning=F,message=F,fig.height=12,fig.width=20}
p1 <- ggarrange(pGP,pCP,labels=c("M9","M10"),font.label = list(size = 20),nrow=1,widths = c(1.2,1))
p2 <- ggarrange(pgpea,prpea,labels=c("M11","M12"),font.label = list(size = 20),nrow=1,widths = c(1.2,1))
fig <- ggarrange(p1,p2,nrow=2)
fig 
if(part=="P1"){
ggsave(fig, file=paste0("../../figs/manuscript/",part,"M9toM12PosteriorDistri.png"),width=20,height=12)  
} else{
ggsave(fig, file=paste0("../../figs/SuppInfo/",part,"M9toM12PosteriorDistri.png"),width=20,height=12)  
}
 
```


