---
title: "Posterior Predictive Checks and Residuals - P2 partition"
author: "Juliette Archambeau"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 4
    toc_float:
       collapsed: false
    number_sections: true
    highlight: textmate
---

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<style type="text/css">
div.main-container {
  max-width: 2000px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 600px;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE,echo=F)
options(width = 300)
library(knitr)
#library(rethinking)
#library(readr)
#library(broom)
library(latex2exp)
library(cowplot)
library(ggplot2)
library(ggpubr)
library(stringi)
library(dplyr)
#library(ggthemes)
library(bayesplot)
library(devtools)
#library(ggbiplot)
#library(arm)
#library(Matrix)
#library(matrixcalc)
library(ggridges)
library(tidyverse)
library(tibble)
library(brms)
#library(kableExtra)
#library(compare)
library(tidybayes)
```



# P2 partition

```{r PartP2, echo=T}
# Data partition used to fit the models
part <- "P2"

# Path
path= paste0("outputs/models/",part,"/")
```

```{r LoadModels}
# Models in a list
myFiles <- list.files(path=paste0("../../",path),pattern=".rds")
models <- list()
for (i in 1:length(myFiles)){
  models[[i]] <- readRDS(file=paste0("../../",path,myFiles[i]))
}
names(models) <- str_sub(myFiles,0,-5)
```

```{r loadTest}
# Load Test dataset
if (grepl("P1",path)==TRUE){
  test <- readRDS(file="../../data/TestP1prepared.RDS")
  train <- readRDS(file="../../data/TrainP1.RDS")
} else if (grepl("P2",path)==TRUE){
  test <- readRDS(file="../../data/TestP2prepared.RDS")
  train <- readRDS(file="../../data/TrainP2.RDS")
}
```

## Some posterior predictive checks

### PPC Density overlay (on the log scale)

Comparison of the empirical distribution of the data $y$ (dark line on the graphs) to the distributions of simulated/replicated data `yrep` from the posterior predictive distribution.


```{r PPCDensOverlay}
figs.train <- list()
figs.test <- list()

for (i in 1:length(models)){
y <- models[[i]]$data$`log(height)`
yrep_train <- posterior_predict(models[[i]], nsamples = 20)

y_test <- log(test$height)
yrep_test <- posterior_predict(models[[i]],newdata=test, nsamples = 20,allow_new_levels = TRUE)

figs.train[[i]] <- pp_check(y,yrep_train,ppc_dens_overlay) + ggtitle(paste0("M",str_sub(names(models)[i],4,-1))) + theme_bw() +
    ggplot2::ylab("Density") + ggplot2::xlab("Log(Height)") +
    theme(plot.title = element_text(size = 20),axis.text=element_text(size=18),
          axis.title = element_text( size = 18), legend.position = c(0.15, 0.8),
          legend.text = element_text(size = 14),legend.background = element_rect(colour = "grey"),
          legend.title = element_blank()) 


figs.test[[i]] <- pp_check(log(test$height), yrep_test,ppc_dens_overlay) + ggtitle(paste0("M",str_sub(names(models)[i],4,-1)))+
    ggplot2::ylab("Density") + ggplot2::xlab("Log(Height)") +
    theme_bw() + theme(plot.title = element_text(size = 20),axis.text=element_text(size=18),
          axis.title = element_text( size = 18), legend.position = c(0.15, 0.8),
          legend.text = element_text(size = 14),legend.background = element_rect(colour = "grey"),
          legend.title = element_blank())
} 
```

#### Predictions on the train dataset 

> Train dataset of the P2 partition => 28 provenances

```{r PPCDensOverlayTrainP2,fig.height=12, fig.width=14}
ggarrange(figs.train[[1]] ,figs.train[[2]], 
          figs.train[[3]],figs.train[[4]],
          figs.train[[5]],figs.train[[6]],
          figs.train[[7]],figs.train[[8]],
          figs.train[[9]],nrow=3,ncol=3,font.label = list(size = 20))
```

#### Predictions of the test dataset

> Test dataset of the P2 partition => 6 provenances

```{r PPCDensOverlayTestP2,fig.height=12,  fig.width=14}
ggarrange(figs.test[[1]] ,figs.test[[2]],
          figs.test[[3]],figs.test[[4]], 
          figs.test[[5]],figs.test[[6]],
          figs.test[[7]],figs.test[[8]],
          figs.test[[9]],nrow=3,ncol=3,font.label = list(size = 20))
```


### PPC violin plot


```{r PPCViolin, fig.height=6}
figs.test.site <- list()
figs.test.prov <- list()

for (i in 1:length(models)){
y_test <- log(test$height)
yrep_test <- posterior_predict(models[[i]],newdata=test, nsamples = 20,allow_new_levels = TRUE)

figs.test.site[[i]] <- ppc_violin_grouped(y_test,yrep_test, group=test$site, probs = c(0.1, 0.5, 0.9),
                                          y_draw =  "both",y_alpha=0.3,y_size=0.5) + 
  theme_bw() + 
  ggtitle(paste0("M",str_sub(names(models)[i],4,-1))) +
  theme(plot.title = element_text(size = 20),axis.text=element_text(size=18),
          axis.title = element_text( size = 18), legend.position = c(0.6, 0.8),
          legend.text = element_text(size = 25),legend.background = element_rect(colour = "grey"),
          legend.title = element_blank()) 

figs.test.prov[[i]] <- ppc_violin_grouped(y_test,yrep_test, group=test$prov, probs = c(0.1, 0.5, 0.9),
                                          y_draw =  "both",y_alpha=0.3,y_size=0.5) + 
  theme_bw() + 
  ggtitle(paste0("M",str_sub(names(models)[i],4,-1))) +
  theme(plot.title = element_text(size = 20),axis.text=element_text(size=18),
          axis.title = element_text( size = 18), 
          legend.text = element_text(size = 25),legend.background = element_rect(colour = "grey"),
          legend.title = element_blank()) 
}
```

#### Predictions in each site

```{r PPCViolinTestSite, fig.width=20,fig.height=12}
ggarrange(figs.test.site[[1]],
          figs.test.site[[2]],
          figs.test.site[[3]],
          figs.test.site[[4]],
          figs.test.site[[5]],
          figs.test.site[[6]],
          figs.test.site[[7]],
          figs.test.site[[8]],
          figs.test.site[[9]],
          nrow=3)
```

#### Predictions in each provenance

```{r PPCViolinTestProv, fig.width=20,fig.height=12}
ggarrange(figs.test.prov[[1]],
          figs.test.prov[[2]],
          figs.test.prov[[3]],
          figs.test.prov[[4]],
          figs.test.prov[[5]],
          figs.test.prov[[6]],
          figs.test.prov[[7]],
          figs.test.prov[[8]],
          figs.test.prov[[9]],
          nrow=3)
```


## Residuals plots

### In the train dataset

```{r ResidualsTrain}
figs.res <- list()

for (i in 1:length(models)){
  
res <- residuals(models[[i]],method="pp_expect",summary=F) %>% t() %>% rowMeans()
fit <- pp_expect(models[[i]]) %>% t() %>% rowMeans()  

figs.res[[i]] <- ggplot(tibble(PredictiveErrors=res,Predictions=fit), aes(x=Predictions, y=PredictiveErrors)) + 
  geom_point(color="gray20",alpha=0.5) + 
  theme_bw() + 
  xlab("Predicted values") +
  ylab("Predive errors") + 
  ggtitle(paste0("M",str_sub(names(models)[i],4,-1))) + 
  theme(axis.text = element_text(size=14),axis.title = element_text(size=14))

}
```

> Predictions against predictive errors

```{r PredictionsVSPredictiveErrorsTrain, fig.width=20, fig.height=14}
ggarrange(figs.res[[1]],
          figs.res[[2]],
          figs.res[[3]],
          figs.res[[4]],
          figs.res[[5]],
          figs.res[[6]],
          figs.res[[7]],
          figs.res[[8]],
          figs.res[[9]],
          nrow=3,ncol=3,font.label = list(size = 20))
```



### In the test data set 

```{r ResidualsTest}
figs.res <- list()

for (i in 1:length(models)){
  
res <- residuals(models[[i]],method="pp_expect",summary=F,newdata=test,allow_new_levels=T) %>% t() %>% rowMeans()
fit <- pp_expect(models[[i]],newdata=test,allow_new_levels=T) %>% t() %>% rowMeans()  

figs.res[[i]] <- ggplot(tibble(PredictiveErrors=res,Predictions=fit), aes(x=Predictions, y=PredictiveErrors)) + 
  geom_point(color="gray20",alpha=0.5) + 
  theme_bw() + 
  xlab("Predicted values") +
  ylab("Predive errors") + 
  ggtitle(paste0("M",str_sub(names(models)[i],4,-1))) + 
  theme(axis.text = element_text(size=14),axis.title = element_text(size=14))

}
```

> Predictions against predictive errors

```{r PredictionsVSPredictiveErrorsTest, fig.width=20, fig.height=14}
ggarrange(figs.res[[1]],
          figs.res[[2]],
          figs.res[[3]],
          figs.res[[4]],
          figs.res[[5]],
          figs.res[[6]],
          figs.res[[7]],
          figs.res[[8]],
          figs.res[[9]],
          nrow=3,ncol=3,font.label = list(size = 20))
```



`r knitr::opts_chunk$set(eval = F)`
