---
title: "Proportion of variance explained and predicted variance conditional on age"
author: "Juliette Archambeau"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    # code_fold: hide
    toc: true
    toc_depth: 4
    toc_float:
       collapsed: false
    number_sections: true
    highlight: textmate
editor_options:
  chunk_output_type: console
---

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<style type="text/css">
div.main-container {
  max-width: 2000px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=F)
library(dplyr)
library(tidyverse)
library(ggpubr)
library(cowplot)
library(stringr)
library(rstanarm)
library(ggforce)
library(tidybayes)
library(latex2exp)
library(ggplot2)
library(kableExtra)
library(brms)
```

# In each common garden

In this section, we aim to calculate $\mathcal{R}^{2}_{ms}|age$ the proportion of the out-of-sample variance (variance of height growth in the test dataset, here 6 left provenances) that is predicted each model $m$ in each common garden $s$, conditional on the age effect.

For that, we first calculate $V_{\text{age}_{2s}}$, the out-of-sample predicted variance by age **in the M2 model** (the baseline model) in each of the common garden, which corresponds to the predicted variance by the fixed effects as there is only the age as fixed effect in this model. 

Then, we calculate $\mathcal{R}^{2}_{ms}|age$ of model $m$ such as:

$$ \mathcal{R}^{2}_{ms}|age = \frac{V_{\text{pred}_{ms}} - V_{\text{age}_{2s}}}{V_{\text{y}_{s}} -V_{\text{age}_{2s}}}$$
where $V_{\text{y}_{s}}$ is the out-of-sample variance in height growth in site $s$ (in the test dataset, i.e. the 6 left provenances) and $V_{\text{pred}_{ms}}$ is the out-of-sample variance predicted by model $m$ in site $s$. 

```{r ChooseBaseline}
# Which baseline model? 
baseline <- "MOD2"
```

## Predictive part


```{r CalculateR2pred}
CalculateR2pred <- function(part){
path= paste0("outputs/models/",part,"/")

# Load the models:
myFiles <- list.files(path=path,pattern=".rds")
models <- list()
for (i in 1:length(myFiles)){
  models[[i]] <- readRDS(file=paste0(path,myFiles[i]))
}
names(models) <- str_sub(myFiles,0,-5)

models <- models[c(baseline,"MOD7","MOD8","MOD9","MOD10","MOD11","MOD12")]


# Load the train and test datasets:
if (grepl("P1",path)==TRUE){
  test <- readRDS(file="data/TestP1prepared.RDS")
  train <- readRDS(file="data/TrainP1.RDS")
} else if (grepl("P2",path)==TRUE){
  test <- readRDS(file="data/TestP2prepared.RDS")
  train <- readRDS(file="data/TrainP2.RDS")
} else if (grepl("P3",path)==TRUE){
  test <- readRDS(file="data/TestP3prepared.RDS")
  train <- readRDS(file="data/TrainP3.RDS")
}

# Site-specific R2
# ================"

# Extract the variance associated with age in each common garden:
var_age_baseline <- lapply(unique(train$site), function(site)
                      pp_expect(models[[baseline]], transform = TRUE,newdata=test[test$site==site,],allow_new_levels=T,re.form=NA))
var_age_baseline <- lapply(var_age_baseline,function(x) apply(x,1,var)) 
var_age_baseline <-lapply(var_age_baseline,mean)
names(var_age_baseline) <- unique(train$site)
# Comment: the variance explained by age is null in Caceres and Madrid as there is only one age!


### Function to extract R2 without age in each site
R2_sites <- function(fit,site) {
  y <- log(test$height[test$site==site])
  var_y <- var(y)
  ypred <- pp_expect(fit, transform = TRUE,newdata=test[test$site==site,],allow_new_levels=T)
  var_ypred <- mean(apply(ypred, 1, var))
  R2 <- (var_ypred - var_age_baseline[[site]]) / (var_y - var_age_baseline[[site]])
}


df <- list()

for (i in unique(train$site)) {
  df[[i]] <- lapply(models, R2_sites,site=i) %>%
    bind_cols(site=i)}


# I don't think it is a good idea to add the global R2 (total variance explained conditional on the age effect) to the graph
# because it includes the variance among sites, so the order of magnitude

# # Global R2
# # ========="
# 
# # Variance of age in M1
# var_age_baseline <-  pp_expect(models$MOD1, transform = TRUE,newdata=test,allow_new_levels=T,re.form=NA) %>% 
#   apply(1,var) %>% 
#   mean
# 
# ### Function to extract R2 without age
# R2 <- function(fit) {
#   y <- log(test$height)
#   var_y <- var(y)
#   ypred <- pp_expect(fit, transform = TRUE,newdata=test,allow_new_levels=T)
#   var_ypred <- mean(apply(ypred, 1, var))
#   R2 <- (var_ypred - var_age_baseline) / (var_y - var_age_baseline)
# }
# 
# df[["All sites"]]  <- lapply(models, R2) %>% bind_cols(site="All sites")

df <- df %>% 
  bind_rows() %>% 
  dplyr::select(site,everything()) %>% 
  mutate(partition=part)
}

df <- lapply(c("P1","P2","P3"),CalculateR2pred)
```


```{r Vizualization1, fig.height=8,fig.width=10, eval=F}
mypalette <- c( "#F46D43","#A6DBA0" ,"#5AAE61","#E0F3F8" , "#ABD9E9", "#74ADD1", "#4575B4")

df[[2]] %>% 
  mutate(site=replace(site, site=="caceres", "Cáceres")) %>% 
  pivot_longer(cols=contains("MOD"),names_to = "model",values_to="var") %>% 
  mutate(model = paste0("M",str_sub(.$model,4,-1)),
         model = fct_relevel(model,c(paste0("M",c(2,7:12)))),
         site=str_to_title(site)) %>% 
  ggplot(aes(fill=model, y=var, x=site)) + 
  geom_bar(position=position_dodge(), stat="identity") +
  theme_bw() +
  labs(fill = "Models") +
  scale_fill_manual(values=mypalette) +
  labs(y=TeX("Proportion of predicted variance $R^{2}_{ms}|age$"),x="") +
  theme(legend.position = c(0.2,0.9),
        axis.text = element_text(size=22),
        axis.title = element_text(size=22), 
        legend.text = element_text(size=18),
        legend.title = element_text(size=20)) +
  guides(fill=guide_legend(ncol=3))
```

```{r Vizualization2, fig.height=12,fig.width=6, eval=F}
df[[2]] %>% 
  mutate(site=replace(site, site=="caceres", "Cáceres")) %>% 
  pivot_longer(cols=contains("MOD"),names_to = "model",values_to="var") %>% 
  mutate(model = paste0("M",str_sub(.$model,4,-1)),
         model = fct_relevel(model,c(paste0("M",c(2,7:12)))),
         site=str_to_title(site)) %>% 
  ggplot(aes(y=var, x=model)) + 
  geom_bar(position=position_dodge(), stat="identity") +
  facet_wrap(~ site, scales = "fixed", ncol = 1) + 
  theme_bw() +
  labs(y=TeX("Proportion of predicted variance $R^{2}_{ms}|age$"),x="") +
  theme(legend.position = c(0.2,0.9),
        axis.text = element_text(size=18),
        axis.title = element_text(size=22), 
        strip.text.x = element_text(size = 20),
        legend.text = element_text(size=18),
        legend.title = element_text(size=20))

df[[2]] %>% 
  mutate(site=replace(site, site=="caceres", "Cáceres")) %>% 
  pivot_longer(cols=contains("MOD"),names_to = "model",values_to="var") %>% 
  mutate(model = paste0("M",str_sub(.$model,4,-1)),
         model = fct_relevel(model,c(paste0("M",c(2,7:12)))),
         site=str_to_title(site)) %>% 
  ggplot(aes(y=var, x=model,fill=model)) + 
  scale_fill_manual(values=mypalette) +
  geom_bar(position=position_dodge(), stat="identity") +
  facet_wrap(~ site, scales = "fixed", ncol = 1) + 
  theme_bw() +
  labs(y=TeX("Proportion of predicted variance $R^{2}_{ms}|age$"),x="") +
  theme(legend.position = "none",
        axis.text = element_text(size=18),
        axis.title = element_text(size=22), 
        strip.text.x = element_text(size = 20),
        legend.text = element_text(size=18),
        legend.title = element_text(size=20))
```

```{r Vizualization3, fig.height=9,fig.width=17, eval=F}
df <- lapply(c("P3"),CalculateR2)

mypalette <- c( "#F46D43","#A6DBA0" ,"#5AAE61","#E0F3F8" , "#ABD9E9", "#74ADD1", "#4575B4")

df[[3]] %>% 
  mutate(site=replace(site, site=="caceres", "Cáceres")) %>% 
  pivot_longer(cols=contains("MOD"),names_to = "model",values_to="var") %>% 
  mutate(model = paste0("M",str_sub(.$model,4,-1)),
         model = fct_relevel(model,c(paste0("M",c(2,7:12)))),
         site=str_to_title(site)) %>% 
  ggplot(aes(fill=model, y=var, x=site)) + 
  geom_bar(position=position_dodge(), stat="identity") +
  theme_bw() +
  labs(fill = "Models") +
  scale_fill_manual(values=mypalette) +
  labs(y=TeX("Proportion of predicted variance $R^{2}_{ms}|age$"),x="") +
  theme(legend.position = c(0.8,0.9),
        axis.text = element_text(size=13),
        axis.title = element_text(size=22), 
        strip.text.x = element_text(size = 25),
        legend.text = element_text(size=18),
        legend.title = element_text(size=20)) +
  guides(fill=guide_legend(ncol=3))


df[[3]] %>% 
  mutate(site=replace(site, site=="caceres", "Cáceres")) %>% 
  pivot_longer(cols=contains("MOD"),names_to = "model",values_to="var") %>% 
  mutate(model = paste0("M",str_sub(.$model,4,-1)),
         model = fct_relevel(model,c(paste0("M",c(2,7:12)))),
         site=str_to_title(site)) %>% 
  ggplot(aes(y=var, x=model,fill=model)) + 
  scale_fill_manual(values=mypalette) +
  geom_bar(position=position_dodge(), stat="identity") +
  facet_grid(.~site,scales="fixed", space = "fixed") + 
  theme_bw() +
  labs(y=TeX("Proportion of predicted variance $R^{2}_{ms}|age$"),x="Models") +
  theme(legend.position = "none",
        axis.text = element_text(size=13),
        strip.text.x = element_text(size = 25),
        axis.title = element_text(size=22), 
        legend.text = element_text(size=18),
        legend.title = element_text(size=20))
```


```{r Visualization4,fig.height=20,fig.width=20}
mypalette <- c( "#F46D43","#A6DBA0" ,"#5AAE61","#E0F3F8" , "#ABD9E9", "#74ADD1", "#4575B4")

p <- df %>% 
  bind_rows() %>% 
  mutate(site=replace(site, site=="caceres", "Cáceres")) %>% 
  #dplyr::select(-MOD14,-MOD1,-MOD0,-MOD3,-MOD4,-MOD5,-MOD6,-MOD13) %>% 
  pivot_longer(cols=contains("MOD"),names_to = "model",values_to="var") %>% 
  mutate(model = paste0("M",str_sub(.$model,4,-1)),
         model = fct_relevel(model,c(paste0("M",str_sub(baseline,4,-1)),paste0("M",c(7:12)))),
         site=str_to_title(site)) %>% 
  ggplot(aes(y=var, x=model,fill=model)) + 
  scale_fill_manual(values=mypalette) +
  geom_bar(position=position_dodge(), stat="identity") +
  facet_grid(partition~site,scales="fixed", space = "fixed") + # ,switch = "y"
  theme_bw() +
  labs(y=TeX("Out-of-sample proportion of predicted variance $R^{2}_{ms}|age$"),x="Models") +
  theme(legend.position = "none",
        axis.text.x = element_text(size=18,angle = 30),
        axis.text.y = element_text(size=25),
        strip.text.x = element_text(size = 25),
        strip.text.y = element_text(size=25,angle = 0),
        axis.title = element_text(size=30), 
        legend.text = element_text(size=18),
        legend.title = element_text(size=20))

p 
# For P1 and P2
#ggsave(p,file="figs/manuscript/ModelComparisonAfterReview.png",width = 20,height = 15)

# For P1, P2 and P3
# ggsave(p,file=paste0("figs/manuscript/ModelComparisonP1P2P3AfterReview_",baseline,".png"),width = 20,height = 20)
```

## Explicative part

```{r CalculateR2explicative}
CalculateR2exp <- function(part){
path= paste0("outputs/models/",part,"/")

# Load the models:
myFiles <- list.files(path=path,pattern=".rds")
models <- list()
for (i in 1:length(myFiles)){
  models[[i]] <- readRDS(file=paste0(path,myFiles[i]))
}
names(models) <- str_sub(myFiles,0,-5)
models <- models[c(baseline,"MOD7","MOD8","MOD9","MOD10","MOD11","MOD12")]

# Load the train and test datasets:
if (grepl("P1",path)==TRUE){
  test <- readRDS(file="data/TestP1prepared.RDS")
  train <- readRDS(file="data/TrainP1.RDS")
} else if (grepl("P2",path)==TRUE){
  test <- readRDS(file="data/TestP2prepared.RDS")
  train <- readRDS(file="data/TrainP2.RDS")
} else if (grepl("P3",path)==TRUE){
  test <- readRDS(file="data/TestP3prepared.RDS")
  train <- readRDS(file="data/TrainP3.RDS")
}

# Site-specific R2
# ================"

# Extract the variance associated with age in each common garden:
var_age_baseline <- lapply(unique(train$site), function(site)
                      pp_expect(models[[baseline]], transform = TRUE,newdata=models[[baseline]]$data[models[[baseline]]$data$site==site,],re.form=NA))
var_age_baseline <- lapply(var_age_baseline,function(x) apply(x,1,var)) 
var_age_baseline <-lapply(var_age_baseline,mean)
names(var_age_baseline) <- unique(train$site)
# Comment: the variance explained by age is null in Caceres and Madrid as there is only one age!


### Function to extract R2 without age in each site
R2_sites <- function(fit,site) {
  y <- fit$data$`log(height)`[fit$data$site==site]
  var_y <- var(y)
  ypred <- pp_expect(fit, transform = TRUE,newdata=fit$data[fit$data$site==site,])
  var_ypred <- mean(apply(ypred, 1, var))
  R2 <- (var_ypred - var_age_baseline[[site]]) / (var_y - var_age_baseline[[site]])
}


df <- list()

for (i in unique(train$site)) {
  df[[i]] <- lapply(models, R2_sites,site=i) %>%
    bind_cols(site=i)}


df <- df %>% 
  bind_rows() %>% 
  dplyr::select(site,everything()) %>% 
  mutate(partition=part)
}

df <- lapply(c("P1","P2","P3"),CalculateR2exp)
```


```{r VisualizationExplicativePart,fig.height=15,fig.width=20}
mypalette <- c( "#F46D43","#A6DBA0" ,"#5AAE61","#E0F3F8" , "#ABD9E9", "#74ADD1", "#4575B4")

p <- df %>% 
  bind_rows() %>% 
  mutate(site=replace(site, site=="caceres", "Cáceres")) %>% 
  pivot_longer(cols=contains("MOD"),names_to = "model",values_to="var") %>% 
  mutate(model = paste0("M",str_sub(.$model,4,-1)),
         model = fct_relevel(model,c(paste0("M",str_sub(baseline,4,-1)),paste0("M",c(7:12)))),
         site=str_to_title(site)) %>% 
  ggplot(aes(y=var, x=model,fill=model)) + 
  scale_fill_manual(values=mypalette) +
  geom_bar(position=position_dodge(), stat="identity") +
  facet_grid(partition~site,scales="fixed", space = "fixed") + # ,switch = "y"
  theme_bw() +
  labs(y=TeX("In-sample proportion of explained variance ($R^{2}_{ms}|age$)"),x="Models") +
  theme(legend.position = "none",
        axis.text.x = element_text(size=18,angle = 30),
        axis.text.y = element_text(size=25),
        strip.text.x = element_text(size = 25),
        strip.text.y = element_text(size=25,angle = 0),
        axis.title = element_text(size=30), 
        legend.text = element_text(size=18),
        legend.title = element_text(size=20))


p 
# For P1 and P2
#ggsave(p,file="figs/manuscript/ModelComparisonAfterReview.png",width = 20,height = 15)

# For P1, P2 and P3
# ggsave(p,file=paste0("figs/SuppInfo/ModelComparison_ExplicativePart_AfterReview_",baseline,".png"),width = 20,height = 20)
```

# In each provenance

In this section, we aim to calculate $\mathcal{R}^{2}_{mp}|age$ the proportion of the out-of-sample variance (variance of height growth in the test dataset, here 6 left provenances) that is predicted each model $m$ for each provenance $p$, conditional on the age effect.

For that, we first calculate $V_{\text{age}_{1p}}$, the out-of-sample predicted variance by age **in the M1 model** in each of the common garden, which corresponds to the predicted variance by the fixed effects as there is only the age as fixed effect in this model. 

Then, we calculate $\mathcal{R}^{2}_{mp}|age$ of model $m$ such as:

$$ \mathcal{R}^{2}_{mp}|age = \frac{V_{\text{pred}_{mp}} - V_{\text{age}_{2p}}}{V_{\text{y}_{p}} -V_{\text{age}_{2p}}}$$
where $V_{\text{y}_{p}}$ is the out-of-sample variance in height growth in the provenance $p$ (in the test dataset) and $V_{\text{pred}_{mp}}$ is the out-of-sample variance predicted by model $m$ in provenance $p$. 


## Predictive part

```{r CalculateR2predProv}
CalculateR2pred <- function(part){
path= paste0("outputs/models/",part,"/")

# Load the models:
myFiles <- list.files(path=path,pattern=".rds")
models <- list()
for (i in 1:length(myFiles)){
  models[[i]] <- readRDS(file=paste0(path,myFiles[i]))
}
names(models) <- str_sub(myFiles,0,-5)

models <- models[c(baseline,"MOD7","MOD8","MOD9","MOD10","MOD11","MOD12")]


# Load the train and test datasets:
if (grepl("P1",path)==TRUE){
  test <- readRDS(file="data/TestP1prepared.RDS")
  train <- readRDS(file="data/TrainP1.RDS")
} else if (grepl("P2",path)==TRUE){
  test <- readRDS(file="data/TestP2prepared.RDS")
  train <- readRDS(file="data/TrainP2.RDS")
} else if (grepl("P3",path)==TRUE){
  test <- readRDS(file="data/TestP3prepared.RDS")
  train <- readRDS(file="data/TrainP3.RDS")
}

# Provenance-specific R2
# ================"

# Extract the variance associated with age in each common garden:
var_age_baseline <- lapply(unique(test$prov), function(prov)
                      pp_expect(models[[baseline]], transform = TRUE,newdata=test[test$prov==prov,],allow_new_levels=T,re.form=NA))
var_age_baseline <- lapply(var_age_baseline,function(x) apply(x,1,var)) 
var_age_baseline <-lapply(var_age_baseline,mean)
names(var_age_baseline) <- unique(test$prov)

### Function to extract R2 without age in each provenance
R2_prov <- function(fit,prov) {
  y <- log(test$height[test$prov==prov])
  var_y <- var(y)
  ypred <- pp_expect(fit, transform = TRUE,newdata=test[test$prov==prov,],allow_new_levels=T)
  var_ypred <- mean(apply(ypred, 1, var))
  R2 <- (var_ypred - var_age_baseline[[prov]]) / (var_y - var_age_baseline[[prov]])
}


df <- list()

for (i in unique(test$prov)) {
  df[[i]] <- lapply(models, R2_prov,prov=i) %>%
    bind_cols(prov=i)}


df <- df %>% 
  bind_rows() %>% 
  dplyr::select(prov,everything()) %>% 
  mutate(partition=part)
}
```


```{r VizualizationPropVarProv, fig.height=8,fig.width=16}
df <- CalculateR2pred("P2")

p2 <- df %>% 
  pivot_longer(cols=contains("MOD"),names_to = "model",values_to="var") %>% 
  mutate(model = paste0("M",str_sub(.$model,4,-1)),
         model = fct_relevel(model,c(paste0("M",str_sub(baseline,4,-1)),paste0("M",c(7:12))))) %>% 
  ggplot(aes(y=var, x=model,fill=model)) + 
  scale_fill_manual(values=mypalette) +
  geom_bar(position=position_dodge(), stat="identity") +
  facet_grid(.~prov,scales="fixed", space = "fixed") + 
  theme_bw() +
  labs(y=TeX("Proportion of predicted variance $R^{2}_{mp}|age$"),x="Models") +
  theme(legend.position = "none",
        axis.text = element_text(size=13),
        strip.text.x = element_text(size = 25),
        axis.title = element_text(size=22), 
        legend.text = element_text(size=18),
        legend.title = element_text(size=20))

p2 


df <- CalculateR2pred("P3")

p3 <- df %>% 
  pivot_longer(cols=contains("MOD"),names_to = "model",values_to="var") %>% 
  mutate(model = paste0("M",str_sub(.$model,4,-1)),
         model = fct_relevel(model,c(paste0("M",str_sub(baseline,4,-1)),paste0("M",c(7:12))))) %>% 
  ggplot(aes(y=var, x=model,fill=model)) + 
  scale_fill_manual(values=mypalette) +
  geom_bar(position=position_dodge(), stat="identity") +
  facet_grid(.~prov,scales="fixed", space = "fixed") + 
  theme_bw() +
  labs(y=TeX("Proportion of predicted variance $R^{2}_{mp}|age$"),x="Models") +
  theme(legend.position = "none",
        axis.text = element_text(size=13),
        strip.text.x = element_text(size = 25),
        axis.title = element_text(size=22), 
        legend.text = element_text(size=18),
        legend.title = element_text(size=20))

p3
```

