---
title: "Qst-Fst analysis"
author: "Juliette Archambeau"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    # code_fold: hide
    toc: true
    toc_depth: 4
    toc_float:
       collapsed: false
    number_sections: true
    highlight: textmate
editor_options:
  chunk_output_type: console
---

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<style type="text/css">
div.main-container {
  max-width: 2000px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=F)
library(dplyr)
library(tidyverse)
library(ggpubr)
library(cowplot)
library(stringr)
library(rstanarm)
library(ggforce)
library(tidybayes)
library(latex2exp)
library(ggplot2)
library(kableExtra)
library(brms)
```

In de Miguel (2021), the global $F_{ST}$ was 0.112 (95% confidence interval: 0.090 - 0.141). According to de Miguel (2021), it "was estimated according to Weir and Cockerham (1984) using the 5,165 SNPs from the Illumina Infinium SNP array and the diveRsity R package (Keenan et al. 2013). The 95% confidence interval of the global $F_{ST}$ estimate was computed by bootstrapping across loci (1,000 bootstrap iterations)."

To calculate the $Q_{ST}$, we use the same formula as in de Miguel (2021), which follows the Spitze (1993), such as:

$$Q_{ST} = \frac{\sigma^2_{P}}{\sigma^2_{P} + 2\sigma^2_{G}} $$

where $\sigma^2_{P}$ is the variance among provenances, and $\sigma^2_{G}$ is the variance among clones (.i.e. genotypes) within provenances.

Quantitative ($Q_{ST}$) and molecular ($F_{ST}$) genetic differentiation among provenances were considered significantly different when $Q_{ST}$ and $F_{ST}$ posterior distributions had non-overlapping 95% confidence intervals.

```{r CalculateQst}
# Load the M1 model fitted on the P1 partition:
mod <- readRDS(file="outputs/models/P1/MOD1.rds")

# Extract the standard deviation of the provenances and clones
sd_prov <- as.array(mod, pars = c("sd_prov__Intercept"), fixed = TRUE) # provenances
sd_clon <- as.array(mod, pars = c("sd_prov:clon__Intercept"), fixed = TRUE) # clones

# Get the variances
sigma_prov <- sd_prov*sd_prov
sigma_clon <- sd_clon*sd_clon

# Calculate the Qst
qst <- sigma_prov / (sigma_prov + 2 * sigma_clon)

# Extract the 95% confidence interval
prob=0.95
probs <- c((1 - prob) / 2, 1 - (1 - prob) / 2)
quantile(qst, probs = probs)

# Extract the median
median(qst)
```


```{r echo=F}
low95 <- round(quantile(qst, probs = probs)[[1]],3)
high95 <- round(quantile(qst, probs = probs)[[2]],3)
med <- round(median(qst),3)
```

As $F_{ST}$ = 0.112 [0.090-0.141] has no overlapping interval with $Q_{ST}$ = `r med` [`r low95`-`r high95`], we conclude that they are significantly different, and therefore that there is adaptive differentiation in height growth in our study.
