---
title: "Calculating counts of gPEAs and rPEAs"
author: "Juliette Archambeau"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    # code_fold: hide
    toc: true
    toc_depth: 4
    toc_float:
       collapsed: false
    number_sections: true
    highlight: textmate
bibliography: references.bib 
---

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<style type="text/css">
div.main-container {
  max-width: 2000px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 600px;
}
```

```{r setup, include=FALSE}
options(width = 300)
library(knitr)
library(readr)
library(dplyr)
library(rethinking)
library(compare)
knitr::opts_chunk$set(cache=F)
```

In this document, we calculate the count of global and regional height-associated positive-effect alleles each genotype in each environment. @miguelPolygenicAdaptationNegative2020 showed that about 350 SNPs were associated with height in maritime pine. Knowing that, we selected gPEAs and rPEAs in two ways:

  - We selected all SNPs with an absolute posterior effect size larger than .006. This leads to the selection of about 350 SNPs.
  
  - **We selected the first 350 SNPs with the largest absolute posterior effect size. This is the method that was adopted for the manuscript.**
  
For selected SNPs with a negative effect on height, we inverted the alleles, so that we only considered alleles with a positive effect on height when calcultating the counts. 

> Load data

```{r  LoadFiles}
# File with the genotype names (clone names)
geno_names <- read.delim2("../../data/ClonapinBlups523IndPiMASSJuly2019.txt", row.names=1)

# File with the genotype of each clone for each SNP
geno <- read.csv("../../data/5165snps523genotypesNA.txt", header=FALSE, row.names=1)
str(geno[,1:20])
dim(geno) # SNPs in rows, genotypes in columns

# In this file, SNPs have their names, but not the genotypes.
head(geno[,1:10])

# Removing the first two columns with allele info (A,T, G or C)
geno <- geno[,3:dim(geno)[[2]]]

# Give the genotype name for each column of geno
colnames(geno) <- rownames(geno_names)

rm(geno_names)

head(geno[,1:10])
dim(geno)
```

There are 523 genotypes and 5,165 SNPs.

**Looking for NAs**

```{r NAgeno}
# Total number of NAs
sum(is.na(geno))

# Nas per rows (Number of NAs per SNPs)
na_row <- rowSums(is.na(geno))
range(na_row)
mean(na_row)

# NAs per columns (Number of NAs per genotypes)
na_col <- colSums(is.na(geno))
range(na_col)
mean(na_col)
```



# Calculating the counts of gPEAs

## Visualizing piMASS outupts

```{r LoadPimassOutputsAllSites, message=F}
beta.snp <- read_delim("../../data/piMASSoutputs/height_all_sites_res.mcmc.txt", "\t", escape_double = FALSE, trim_ws = TRUE)
beta.snp <- as.data.frame(beta.snp)
row.names(beta.snp) <- beta.snp$rs
beta.snp$rs <- NULL
head(beta.snp)

# Checking that the genotypes names in the two files are the same,
# and are ranked similarly.
compare::compare(row.names(beta.snp),row.names(geno))
```


```{r VisualizePimassOutputsAllSites, fig.height=8,fig.width=10,echo=F}
par(mfrow=c(2,2))
hist(beta.snp$postc,100L,main="Estimates of the posterior inclusion\nprobabilities based on simple counting")
hist(beta.snp$postrb,100L,main="Estimates of the posterior inclusion\nprobabilities based on Rao-Blackwellization")
hist(beta.snp$beta,100L,main="Naive estimates of the\nposterior effect size")
hist(beta.snp$betarb,100L,main="Rao-Blackwellized estimates\nof the posterior effect size")
```

> Posterior inclusion probabilities

```{r VisualizingPimassOutputsHighpostrbAllSites01, fig.height=4,fig.width=6,echo=F}
hist(beta.snp$postrb[beta.snp$postrb>0.1],100L,main="Estimates of the posterior inclusion probabilities\nbased on Rao-Blackwellization\nwith only the SNPs with postrb>0.1")
```

Below are the estimates of the posterior inclusion probabilities based on Rao-Blackwellization with the four SNPs with **postrb>0.4**.

```{r VisualizingPimassOutputsHighpostrbAllSites04}
beta.snp[abs(beta.snp$postrb)>0.4|abs(beta.snp$postrb)==0.4,]
```

> Posterior effect sizes

```{r VisualizingPimassOutputsHighbetarbAllSites01, fig.height=4,fig.width=6,echo=F}
hist(beta.snp$betarb[abs(beta.snp$betarb)>0.01],100L,main="Rao-Blackwellized estimates\nof the posterior effect size\nwith only the SNPs with betarb>0.01")
```

Below are the Rao-Blackwellized estimates the posterior effect size with the three SNPs with **betarb>0.05**.

```{r VisualizingPimassOutputsHighbetarbAllSites04}
beta.snp[abs(beta.snp$betarb)>0.05|abs(beta.snp$betarb)==0.05,]
```


## Selecting gPEAs (with betarb>0.006)

```{r SNPHeightAssociatedAllSites}
# SNPs associated with height (with posterior effect size higher than 0.006)
snps.h <- beta.snp[abs(beta.snp$betarb)>0.006|abs(beta.snp$betarb)==0.006,]
length(snps.h$chr)

# Among the selected SNPs, SNPs positively associated with height:
snps.h.pos <- row.names(snps.h[snps.h$betarb>0,])
length(snps.h.pos)

# Among the selected SNPs, SNPs negatively associated with height:
snps.h.neg <- row.names(snps.h[snps.h$betarb<0,])
length(snps.h.neg)
```

```{r CountPositiveSNPsAllSites,fig.height=3,fig.width=3}
# Subset of alleles with positive effect on height
geno.snp.h.pos <- geno[row.names(geno) %in% snps.h.pos,]
geno.snp.h.pos[1:5,1:5]

# Subset of alleles with negative effect on height
geno.snp.h.neg <- geno[row.names(geno) %in% snps.h.neg,]
geno.snp.h.neg[1:5,1:5]

# Inverting 0 and 2 in the subset of alleles with negative effects, 
# to only have alleles with a positive effect
snpnames <- row.names(geno.snp.h.neg)
geno.snp.h.neg <- geno.snp.h.neg %>%  mutate_if(is.integer, as.numeric) %>% mutate_all(list(~recode(., `0` = 2,`2`=0)))
row.names(geno.snp.h.neg) <- snpnames 
geno.snp.h.neg[1:5,1:5]

# combining positive and negative alleles
geno.snp.h <-rbind(geno.snp.h.pos,geno.snp.h.neg)
geno.snp.h[1:5,1:5]

DF <- data.frame(clon=names(geno.snp.h),count_all_006=apply(geno.snp.h,2,sum,na.rm=T))

var(DF$count_all_006)
range(DF$count_all_006)
mean(DF$count_all_006)
rethinking::dens(DF$count_all_006)
```


## Selecting gPEAs (first 350 SNPs)

```{r SnpHeightAssociated350AllSites}
# Snps associated with height
beta.snp.order <- beta.snp[order(abs(beta.snp$betarb),decreasing=T),]
beta.snp.order350 <- beta.snp.order[1:350,]
length(beta.snp.order350$beta)

# Snps positively associated with height
snps.h.pos.350 <- row.names(beta.snp.order350[beta.snp.order350$betarb>0,])
length(snps.h.pos.350)

# Snps negatively associated with height
snps.h.neg.350 <- row.names(beta.snp.order350[beta.snp.order350$betarb<0,])
length(snps.h.neg.350)
```

```{r CountPositive350snpsAllSites,fig.height=3,fig.width=3}
# Subset of alleles with positive effect on height
geno.snp.h.pos.350 <- geno[row.names(geno) %in% snps.h.pos.350,]
geno.snp.h.pos.350[1:5,1:5]

# Subset of alleles with negative effect on height
geno.snp.h.neg.350 <- geno[row.names(geno) %in% snps.h.neg.350,]
geno.snp.h.neg.350[1:5,1:5]

# Inverting 0 and 2 in the subset of alleles with negative effects, 
# to only have alleles with a positive effect
snpnames <- row.names(geno.snp.h.neg.350)
geno.snp.h.neg.350 <- geno.snp.h.neg.350 %>%  mutate_if(is.integer, as.numeric) %>% mutate_all(list(~recode(., `0` = 2,`2`=0)))
row.names(geno.snp.h.neg.350) <- snpnames 
geno.snp.h.neg.350[1:5,1:5]

# combining positive and negative alleles
geno.snp.h.350 <-rbind(geno.snp.h.pos.350,geno.snp.h.neg.350)
names.all <- rownames(geno.snp.h.350)

df <- data.frame(clon=names(geno.snp.h.350),count_all_350=apply(geno.snp.h.350,2,sum,na.rm=T))

var(df$count_all_350)
range(df$count_all_350)
mean(df$count_all_350)
rethinking::dens(df$count_all_350)

DF <- merge(DF,df,by="clon")
head(DF)
```


# Calculate the regional counts of rPEAs

## In the French Atlantic region

### Visualizing piMASS outupts

```{r LoadPimassOutputsFrenchAtlantic, message=F}
beta.snp <- read_delim("../../../data/raw-data/res_analyses-max2000_31072019/height_french_atlantic_res.mcmc.txt", "\t", escape_double = FALSE, trim_ws = TRUE)
beta.snp <- as.data.frame(beta.snp)
row.names(beta.snp) <- beta.snp$rs
beta.snp$rs <- NULL
head(beta.snp)
```

```{r VisualizingPimassOutputsFrenchAtlantic, fig.height=8,fig.width=10,echo=F}
par(mfrow=c(2,2))
hist(beta.snp$postc,100L,main="Estimates of the posterior inclusion\nprobabilities based on simple counting")
hist(beta.snp$postrb,100L,main="Estimates of the posterior inclusion\nprobabilities based on Rao-Blackwellization")
hist(beta.snp$beta,100L,main="Naive estimates of the\nposterior effect size")
hist(beta.snp$betarb,100L,main="Rao-Blackwellized estimates\nof the posterior effect size")
```


> Posterior inclusion probabilities

```{r VisualizePimasOutputsHighpostrbFrenchAtlantic01, fig.height=4,fig.width=6,echo=F}
hist(beta.snp$postrb[beta.snp$postrb>0.1],100L,main="Estimates of the posterior inclusion probabilities\nbased on Rao-Blackwellization\nwith only the SNPs with postrb>0.1")
```

Below are the estimates of the posterior inclusion probabilities based on Rao-Blackwellization with the three SNPs with **postrb>0.4**.


```{r VisualizePimasOutputsHighpostrbFrenchAtlantic04}
beta.snp[abs(beta.snp$postrb)>0.4|abs(beta.snp$postrb)==0.4,]
```

> Posterior effect sizes

```{r VisualizePimassOutputsHighbetarbFrenchAtlantic001, fig.height=4,fig.width=6,echo=F}
hist(beta.snp$betarb[abs(beta.snp$betarb)>0.01],100L,main="Rao-Blackwellized estimates\nof the posterior effect size\nwith only the SNPs with betarb>0.01")
```

Below are the Rao-Blackwellized estimates the posterior effect size with the two SNPs with **betarb>0.05**.

```{r VisualizePimassOutputsHighbetarbFrenchAtlantic005}
beta.snp[abs(beta.snp$betarb)>0.05|abs(beta.snp$betarb)==0.05,]
```



### Selecting rPEAs (betarb>0.006)


```{r SnpHeightAssociatedFrenchAtlantic}
# Snps associated with height
snps.h <- beta.snp[abs(beta.snp$betarb)>0.006|abs(beta.snp$betarb)==0.006,]
length(snps.h$chr)

# Snps positively associated with height
snps.h.pos <- row.names(snps.h[snps.h$betarb>0,])
length(snps.h.pos)

# Snps negatively associated with height
snps.h.neg <- row.names(snps.h[snps.h$betarb<0,])
length(snps.h.neg)
```

```{r CountPositiveSnpsFrenchAtlantic,fig.height=3,fig.width=3}
# Subset of alleles with positive effect on height
geno.snp.h.pos <- geno[row.names(geno) %in% snps.h.pos,]
geno.snp.h.pos[1:5,1:5]

# Subset of alleles with negative effect on height
geno.snp.h.neg <- geno[row.names(geno) %in% snps.h.neg,]
geno.snp.h.neg[1:5,1:5]

# Inverting 0 and 2 in the subset of alleles with negative effects, 
# to only have alleles with a positive effect
snpnames <- row.names(geno.snp.h.neg)
geno.snp.h.neg <- geno.snp.h.neg %>%  mutate_if(is.integer, as.numeric) %>% mutate_all(list(~recode(., `0` = 2,`2`=0)))
row.names(geno.snp.h.neg) <- snpnames 
geno.snp.h.neg[1:5,1:5]

# combining positive and negative alleles
geno.snp.h <-rbind(geno.snp.h.pos,geno.snp.h.neg)

df <- data.frame(clon=names(geno.snp.h),count_fratl_006=apply(geno.snp.h,2,sum,na.rm=T))

var(df$count_fratl_006)
range(df$count_fratl_006)
mean(df$count_fratl_006)
rethinking::dens(df$count_fratl_006)

DF <- merge(DF,df,by="clon")
head(DF)
```


### Selecting rPEAs (first 350 SNPs)

```{r SnpHeightAssociated350FrenchAtlantic}
# Snps associated with height
beta.snp.order <- beta.snp[order(abs(beta.snp$betarb),decreasing=T),]
beta.snp.order350 <- beta.snp.order[1:350,]
length(beta.snp.order350$beta)

# Snps positively associated with height
snps.h.pos.350 <- row.names(beta.snp.order350[beta.snp.order350$betarb>0,])
length(snps.h.pos.350)

# Snps negatively associated with height
snps.h.neg.350 <- row.names(beta.snp.order350[beta.snp.order350$betarb<0,])
length(snps.h.neg.350)
```

```{r CountPositive350snpsFrenchAtlantic,fig.height=3,fig.width=3}
# Subset of alleles with positive effect on height
geno.snp.h.pos.350 <- geno[row.names(geno) %in% snps.h.pos.350,]

# Subset of alleles with negative effect on height
geno.snp.h.neg.350 <- geno[row.names(geno) %in% snps.h.neg.350,]
geno.snp.h.neg.350[1:5,1:5]

# Inverting 0 and 2 in the subset of alleles with negative effects, 
# to only have alleles with a positive effect
snpnames <- row.names(geno.snp.h.neg.350)
geno.snp.h.neg.350 <- geno.snp.h.neg.350 %>%  mutate_if(is.integer, as.numeric) %>% mutate_all(list(~recode(., `0` = 2,`2`=0)))
row.names(geno.snp.h.neg.350) <- snpnames 
geno.snp.h.neg.350[1:5,1:5]

# combining positive and negative alleles
geno.snp.h.350 <-rbind(geno.snp.h.pos.350,geno.snp.h.neg.350)
names.fr_atl <- rownames(geno.snp.h.350)

df <- data.frame(clon=names(geno.snp.h.350),count_fratl_350=apply(geno.snp.h.350,2,sum,na.rm=T))

var(df$count_fratl_350)
range(df$count_fratl_350)
mean(df$count_fratl_350)
rethinking::dens(df$count_fratl_350)

DF <- merge(DF,df,by="clon")
head(DF)
```



## In the Iberian Atlantic region

### Visualizing piMASS outupts

```{r LoadPimassOutputsIberianAtlantic, message=F}
beta.snp <- read_delim("../../../data/raw-data/res_analyses-max2000_31072019/height_iberian_atlantic_res.mcmc.txt", "\t", 
                       escape_double = FALSE, trim_ws = TRUE)
beta.snp <- as.data.frame(beta.snp)
row.names(beta.snp) <- beta.snp$rs
beta.snp$rs <- NULL
head(beta.snp)
```

```{r VisualizingPimassOutputsIberianAtlantic, fig.height=8,fig.width=10,echo=F}
par(mfrow=c(2,2))
hist(beta.snp$postc,100L,main="Estimates of the posterior inclusion\nprobabilities based on simple counting")
hist(beta.snp$postrb,100L,main="Estimates of the posterior inclusion\nprobabilities based on Rao-Blackwellization")
hist(beta.snp$beta,100L,main="Naive estimates of the\nposterior effect size")
hist(beta.snp$betarb,100L,main="Rao-Blackwellized estimates\nof the posterior effect size")
```

> Posterior inclusion probabilities

```{r VisualizePimasOutputsHighpostrbIberianAtlantic01, fig.height=4,fig.width=6,echo=F}
hist(beta.snp$postrb[beta.snp$postrb>0.1],100L,main="Estimates of the posterior inclusion probabilities\nbased on Rao-Blackwellization\nwith only the SNPs with postrb>0.1")
```

Below are the estimates of the posterior inclusion probabilities based on Rao-Blackwellization with the three SNPs with **postrb>0.4**.


```{r VisualizePimasOutputsHighpostrbIberianAtlantic04}
beta.snp[abs(beta.snp$postrb)>0.4|abs(beta.snp$postrb)==0.4,]
```

> Posterior effect sizes

```{r VisualizePimassOutputsHighbetarbIberianAtlantic001, fig.height=4,fig.width=6,echo=F}
hist(beta.snp$betarb[abs(beta.snp$betarb)>0.01],100L,main="Rao-Blackwellized estimates\nof the posterior effect size\nwith only the SNPs with betarb>0.01")
```

Below are the Rao-Blackwellized estimates the posterior effect size with the two SNPs with **betarb>0.05**.

```{r VisualizePimassOutputsHighbetarbIberianAtlantic005}
beta.snp[abs(beta.snp$betarb)>0.05|abs(beta.snp$betarb)==0.05,]
```




### Selecting rPEAS (betarb>0.006)

```{r SnpHeightAssociatedIberianAtlantic}
# Snps associated with height
snps.h <- beta.snp[abs(beta.snp$betarb)>0.006|abs(beta.snp$betarb)==0.006,]
length(snps.h$chr)

# Snps positively associated with height
snps.h.pos <- row.names(snps.h[snps.h$betarb>0,])
length(snps.h.pos)

# Snps negatively associated with height
snps.h.neg <- row.names(snps.h[snps.h$betarb<0,])
length(snps.h.neg)
```

```{r CountPositiveSnpsIberianAtlantic,fig.height=3,fig.width=3}
# Subset of alleles with positive effect on height
geno.snp.h.pos <- geno[row.names(geno) %in% snps.h.pos,]
geno.snp.h.pos[1:5,1:5]

# Subset of alleles with negative effect on height
geno.snp.h.neg <- geno[row.names(geno) %in% snps.h.neg,]
geno.snp.h.neg[1:5,1:5]

# Inverting 0 and 2 in the subset of alleles with negative effects, 
# to only have alleles with a positive effect
snpnames <- row.names(geno.snp.h.neg)
geno.snp.h.neg <- geno.snp.h.neg %>%  mutate_if(is.integer, as.numeric) %>% mutate_all(list(~recode(., `0` = 2,`2`=0)))
row.names(geno.snp.h.neg) <- snpnames 
geno.snp.h.neg[1:5,1:5]

# combining positive and negative alleles
geno.snp.h <-rbind(geno.snp.h.pos,geno.snp.h.neg)

df <- data.frame(clon=names(geno.snp.h),count_ibatl_006=apply(geno.snp.h,2,sum,na.rm=T))

var(df$count_ibatl_006)
range(df$count_ibatl_006)
mean(df$count_ibatl_006)
rethinking::dens(df$count_ibatl_006)

DF <- merge(DF,df,by="clon")
head(DF)
```



### Selecting rPEAs (first 350 SNPs)


```{r SnpHeightAssociated350IberianAtlantic}
# Snps associated with height
beta.snp.order <- beta.snp[order(abs(beta.snp$betarb),decreasing=T),]
beta.snp.order350 <- beta.snp.order[1:350,]
length(beta.snp.order350$beta)

# Snps positively associated with height
snps.h.pos.350 <- row.names(beta.snp.order350[beta.snp.order350$betarb>0,])
length(snps.h.pos.350)

# Snps negatively associated with height
snps.h.neg.350 <- row.names(beta.snp.order350[beta.snp.order350$betarb<0,])
length(snps.h.neg.350)
```

```{r CountPositive350snpsIberianAtlantic,fig.height=3,fig.width=3}
# Subset of alleles with positive effect on height
geno.snp.h.pos.350 <- geno[row.names(geno) %in% snps.h.pos.350,]
geno.snp.h.pos.350[1:5,1:5]

# Subset of alleles with negative effect on height
geno.snp.h.neg.350 <- geno[row.names(geno) %in% snps.h.neg.350,]
geno.snp.h.neg.350[1:5,1:5]

# Inverting 0 and 2 in the subset of alleles with negative effects, 
# to only have alleles with a positive effect
snpnames <- row.names(geno.snp.h.neg.350)
geno.snp.h.neg.350 <- geno.snp.h.neg.350 %>%  mutate_if(is.integer, as.numeric) %>% mutate_all(list(~recode(., `0` = 2,`2`=0)))
row.names(geno.snp.h.neg.350) <- snpnames 
geno.snp.h.neg.350[1:5,1:5]

# combining positive and negative alleles
geno.snp.h.350 <-rbind(geno.snp.h.pos.350,geno.snp.h.neg.350)
names.ib_atl <- rownames(geno.snp.h.350)

df <- data.frame(clon=names(geno.snp.h.350),count_ibatl_350=apply(geno.snp.h.350,2,sum,na.rm=T))

var(df$count_ibatl_350)
range(df$count_ibatl_350)
mean(df$count_ibatl_350)
rethinking::dens(df$count_ibatl_350)

DF <- merge(DF,df,by="clon")
head(DF)
```



## In the Mediterranean region

### Visualizing piMASS outupts

```{r LoadPimassOutputsMed, message=F}
beta.snp <- read_delim("../../../data/raw-data/res_analyses-max2000_31072019/height_mediterranean_res.mcmc.txt", "\t", 
                       escape_double = FALSE, trim_ws = TRUE)
beta.snp <- as.data.frame(beta.snp)
row.names(beta.snp) <- beta.snp$rs
beta.snp$rs <- NULL
head(beta.snp)
```

```{r VisualizingPimassOutputsMed, fig.height=8,fig.width=10,echo=F}
par(mfrow=c(2,2))
hist(beta.snp$postc,100L,main="Estimates of the posterior inclusion\nprobabilities based on simple counting")
hist(beta.snp$postrb,100L,main="Estimates of the posterior inclusion\nprobabilities based on Rao-Blackwellization")
hist(beta.snp$beta,100L,main="Naive estimates of the\nposterior effect size")
hist(beta.snp$betarb,100L,main="Rao-Blackwellized estimates\nof the posterior effect size")
```

> Posterior inclusion probabilities

```{r VisualizePimasOutputsHighpostrbMedAtlantic01, fig.height=4,fig.width=6,echo=F}
hist(beta.snp$postrb[beta.snp$postrb>0.1],100L,
     main="Estimates of the posterior inclusion probabilities\nbased on Rao-Blackwellization\nwith only the SNPs with postrb>0.1")
```

Below are the estimates of the posterior inclusion probabilities based on Rao-Blackwellization with the three SNPs with **postrb>0.4**.

```{r VisualizePimasOutputsHighpostrbMedAtlantic04}
beta.snp[abs(beta.snp$postrb)>0.4|abs(beta.snp$postrb)==0.4,]
```

> Posterior effect sizes

```{r VisualizePimassOutputsHighbetarbMedAtlantic001, fig.height=4,fig.width=6,echo=F}
hist(beta.snp$betarb[abs(beta.snp$betarb)>0.01],100L,
     main="Rao-Blackwellized estimates\nof the posterior effect size\nwith only the SNPs with betarb>0.01")
```

Below are the Rao-Blackwellized estimates the posterior effect size with the two SNPs with **betarb>0.05**.

```{r VisualizePimassOutputsHighbetarbMedAtlantic005}
beta.snp[abs(beta.snp$betarb)>0.05|abs(beta.snp$betarb)==0.05,]
```



### Selecting rPEAS (betarb>0.006)


```{r SnpHeightAssociatedMed}
# Snps associated with height
snps.h <- beta.snp[abs(beta.snp$betarb)>0.006|abs(beta.snp$betarb)==0.006,]
length(snps.h$chr)

# Snps positively associated with height
snps.h.pos <- row.names(snps.h[snps.h$betarb>0,])
length(snps.h.pos)

# Snps negatively associated with height
snps.h.neg <- row.names(snps.h[snps.h$betarb<0,])
length(snps.h.neg)
```

```{r CountPositiveSnpsMed,fig.height=3,fig.width=3}
# Subset of alleles with positive effect on height
geno.snp.h.pos <- geno[row.names(geno) %in% snps.h.pos,]
geno.snp.h.pos[1:5,1:5]

# Subset of alleles with negative effect on height
geno.snp.h.neg <- geno[row.names(geno) %in% snps.h.neg,]
geno.snp.h.neg[1:5,1:5]

# Inverting 0 and 2 in the subset of alleles with negative effects, 
# to only have alleles with a positive effect
snpnames <- row.names(geno.snp.h.neg)
geno.snp.h.neg <- geno.snp.h.neg %>%  mutate_if(is.integer, as.numeric) %>% mutate_all(list(~recode(., `0` = 2,`2`=0)))
row.names(geno.snp.h.neg) <- snpnames 
geno.snp.h.neg[1:5,1:5]

# combining positive and negative alleles
geno.snp.h <-rbind(geno.snp.h.pos,geno.snp.h.neg)

df <- data.frame(clon=names(geno.snp.h),count_med_006=apply(geno.snp.h,2,sum,na.rm=T))

var(df$count_med_006)
range(df$count_med_006)
mean(df$count_med_006)
rethinking::dens(df$count_med_006)

DF <- merge(DF,df,by="clon")
head(DF)
```



### Selecting rPEAs (first 350 SNPs)

```{r SnpHeightAssociated350Med}
# Snps associated with height
beta.snp.order <- beta.snp[order(abs(beta.snp$betarb),decreasing=T),]
beta.snp.order350 <- beta.snp.order[1:350,]
length(beta.snp.order350$beta)

# Snps positively associated with height
snps.h.pos.350 <- row.names(beta.snp.order350[beta.snp.order350$betarb>0,])
length(snps.h.pos.350)

# Snps negatively associated with height
snps.h.neg.350 <- row.names(beta.snp.order350[beta.snp.order350$betarb<0,])
length(snps.h.neg.350)
```

```{r CountPositive350snpsMed,fig.height=3,fig.width=3}
# Subset of alleles with positive effect on height
geno.snp.h.pos.350 <- geno[row.names(geno) %in% snps.h.pos.350,]
geno.snp.h.pos.350[1:5,1:5]

# Subset of alleles with negative effect on height
geno.snp.h.neg.350 <- geno[row.names(geno) %in% snps.h.neg.350,]
geno.snp.h.neg.350[1:5,1:5]

# Inverting 0 and 2 in the subset of alleles with negative effects, 
# to only have alleles with a positive effect
snpnames <- row.names(geno.snp.h.neg.350)
geno.snp.h.neg.350 <- geno.snp.h.neg.350 %>%  mutate_if(is.integer, as.numeric) %>% mutate_all(list(~recode(., `0` = 2,`2`=0)))
row.names(geno.snp.h.neg.350) <- snpnames 
geno.snp.h.neg.350[1:5,1:5]

# combining positive and negative alleles
geno.snp.h.350 <-rbind(geno.snp.h.pos.350,geno.snp.h.neg.350)
names.med <- rownames(geno.snp.h.350)

df <- data.frame(clon=names(geno.snp.h.350),count_med_350=apply(geno.snp.h.350,2,sum,na.rm=T))

var(df$count_med_350)
range(df$count_med_350)
mean(df$count_med_350)
rethinking::dens(df$count_med_350)

DF <- merge(DF,df,by="clon")
DF
dim(DF)
sum(is.na(DF))
```

# Save all

```{r SaveDF,eval=F}
saveRDS(DF,file="../../data/CountPEAs.RDS")
```

# Overlap between the 350 SNPs selected in each region


```{r overlapsrPEAs}
length(setdiff(names.fr_atl,names.ib_atl))
length(setdiff(names.fr_atl,names.med))
length(setdiff(names.ib_atl,names.med))
```

> 69 PEAs shared between Iberian and 192 French Atlantic regions, 41 shared between French Atlantic and Mediterranean regions, and 83 shared between the Iberian Atlantic and Mediterranean regions.

# References


`r knitr::opts_chunk$set(eval = F)`




