---
title: "Climatic similarity matrices"
author: "Juliette Archambeau"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    # code_fold: hide
    toc: true
    toc_depth: 4
    toc_float:
       collapsed: false
    number_sections: true
    highlight: textmate
bibliography: references.bib 
---

<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

<style type="text/css">
div.main-container {
  max-width: 2000px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{css, echo=FALSE}
pre {
  max-height: 300px;
  overflow-y: auto;
}

pre[class] {
  max-height: 600px;
}
```

```{r setup, include=FALSE}
options(width = 300)
library(knitr)
library(readr)
library(cowplot)
library(ggplot2)
library(dplyr)
library(gplots)
library(usdm)
library(devtools)
library(ggbiplot)
library(arm)
library(Matrix)
library(matrixcalc)
library(compare)
library(RColorBrewer)
library(MASS)
knitr::opts_chunk$set(fig.width = 12,cache=FALSE)
options(width = 150)
```

<!-- Functions used -->

```{r FuntionsUsed,echo=F}
source("../../scripts/Functions/corpmat.R")
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))
```

# Climatic similarity between test sites (P1)

## Load and clean data


```{r LoadDataSites}
data <- readRDS(file="../../data/TrainP1.RDS")
```

Let's create a column to identify unique site + year of measurement (= different age of the trees).

```{r CreateColumnSiteAgeSites}
data$site_age <- paste0(data$site,data$age)
```

> Keeping the columns of interest

- Some variables have no biological interest, we can remove them: `tmx_min_1y_site` and `tmn_max_1y_site`

- Minimum precipitation in Portugal are equal to 1.2mm for three measurments. But it corresponds to precipitation in winter (in March if I well remembered). And trees are in dormancy in winter, and are not affected (or at least less) by the lack of water. So better to use other variable like `pre_summer_min_site` which corresponds to the precipitation during the previous summer. 

```{r KeepColumnOfInterestSites}
data <- data[,c("site_age",grep("y_site",colnames(data),value=T),grep("summer_min",colnames(data),value=T))]

# Removing of variables with no biologcial interest
data[,c("tmx_min_1y_site","tmn_max_1y_site")] <- NULL

# Very weak precipitation in Portugal in winter. Let's remove this variable.
table(data$pre_min_1y_site,data$site_age)
data[,c("pre_min_1y_site")] <- NULL
```

Keep only unique values:

```{r UniqueValuesSites}
data <- unique(data)
```


## Exploratory analyses


> Checking NAs

```{r noNAsSites}
sapply(data, function(x) sum(is.na(x)))
```


> Mean-centered variables

```{r ScalingVariablesSites}
data[,paste0(grep("_site",names(data),value = T),".sc")] <- scale(data[,grep("_site",names(data),value = T)])
```


```{r SquareCorrplotDataSites, fig.height=12}
cor <- cor(data[,grep("_site.sc",names(data),value = T)])
corrplot::corrplot(cor,method="number")
```


```{r CorrplotTriangleSites, fig.height=12}
# matrix of the p-value of the correlation
p.mat <- corpmat(cor)

corrplot::corrplot(cor, method="color", col=col(200),  
                   type="upper", order="hclust", 
                   addCoef.col = "black", # Add coefficient of correlation
                   tl.col="black", tl.srt=45, #Text label color and rotation
                   # Combine with significance
                   p.mat = p.mat, sig.level = 0.01, insig = "blank", 
                   # hide correlation coefficient on the principal diagonal
                   diag=FALSE)
```

```{r PCASites, fig.height=12}
pca <- prcomp(data[,grep("_site.sc",names(data),value = T)], center = TRUE,scale. = TRUE)
ggbiplot(pca,varname.size =4) +  ylim(-4.5, 2.5) +    xlim(-3, 3) +  
   theme_minimal(base_size = 25) +
  theme(plot.title = element_text(size=18))
```

## Selected variables

Based on the exploratory analyses, we selected 6 variables describing the climatic conditions between test test sites (3 precipitation-related variables and 3 temperature-related variables). 

- `pre_summer_min_site` (*min.presummer* in the manuscript): the minimum of the monthly precipitation during summer -June to September- (°C)

- `pre_mean_1y_site` (*mean.pre* in the manuscript): the mean of the monthly precipitation (mm)

- `pre_max_1y_site` (*max.pre* in the manuscript): the maximum of the monthly precipitation ( mm)

- `tmn_min_1y_site` (*min.tmn* in the manuscript):  the minimum of the monthly minimum temperatures (°C)

- `tmx_mean_1y_site.sc` (*mean.tmax* in the manuscript): the mean of monthly maximum temperatures (°C)

- `tmx_max_1y_site` (*max.tmx* in the manuscript): the maximum of the monthly maximum temperatures (°C)


```{r SelectedVariablesSites}
select.var <- c("pre_summer_min_site.sc","pre_mean_1y_site.sc","tmn_min_1y_site.sc",
                "tmx_max_1y_site.sc","pre_max_1y_site.sc","tmx_mean_1y_site.sc")
```

```{r CorrelationsSites}
cor(data[,select.var])
```

```{r VIFSites}
vif(data[,select.var])
```

## Matrices of environmental similarity

2 options:

- A matrix of Euclidean distance following @thomsonGuideUsingMultiplematrix2018.
    
- A variance-covariance matrix following @jarquinReactionNormModel2014.


```{r PrepareDatasetSites}
dat <- unique(data[,c("site_age",select.var)])
row.names(dat) <- dat$site_age
dat$site_age <- NULL 
round(dat,3)
```

### Matrix of Euclidean distance 

Following @thomsonGuideUsingMultiplematrix2018.

I did not select this option in the manuscript as I did not know how to include a euclidean matrix in `brms`. 

```{r EucliMatrixSites}
euclimat <-as.matrix(dist(dat, method="euclidean", diag=TRUE,upper=TRUE))
round(euclimat,3)

# then scales so that the values are between 0 and 1
euclimat <- 1- euclimat/max(euclimat, na.rm=TRUE)
round(euclimat,3)
```

```{r SaveEucliMatrixSites}
# saveRDS(euclimat, file = "../../data/EnvMat/EucliMatrixSites.rds")
```


### Variance-covariance matrix

Following @jarquinReactionNormModel2014.

This is the option used in the mannuscript.

```{r VarCovMatrixSites, fig.height=8,fig.width=6}
varmat <-as.matrix(var(t(dat)))
round(varmat,3)
heatmap(varmat)

heatmap.2(as.matrix(varmat),scale="none",
          col=brewer.pal(11,"RdBu"),
          trace="none",key=F,
          cexRow=1.5,cexCol=1.5,
          margins=c(12,8))
```

```{r SaveHeatmapVarCovMatrixSites, echo=F,message=F}
png("../../figs/SuppInfo/ExploringData/VarCovMatrixSites.png",width = 800,height=800)
heatmap.2(as.matrix(varmat),scale="none",
          col=brewer.pal(11,"RdBu"),
          trace="none",key=F,
          cexRow=1.5,cexCol=1.5,
          margins=c(12,8))
dev.off() 
```

```{r PosDefVarCovMatrixSites}
is.symmetric.matrix(varmat)
is.positive.definite(varmat)
varmat.pd <- nearPD(varmat)$mat
all.equal(as.matrix(varmat.pd),varmat)
varmat.pd
```


```{r SavePosDefVarCovMatrixSites}
# saveRDS(varmat.pd, file = "../../data/cleaned-data/EnvMat/VarCovMatSites.rds")
```



# Climatic similarity between provenances (P1)

## Load and clean data


```{r LoadDataProvenancesP1}
data <- readRDS(file="../../data/TrainP1.RDS")
```


> Keeping the columns of interest

```{r KeepColumnOfInterestProvenancesP1}
data <- data[,c("prov",grep("_prov",colnames(data),value=T))]
data <- data[,-(grep("tude",colnames(data)))]

# Removing soil variables
data <- data[,-(grep("_sub_",colnames(data)))]
data <- data[,-(grep("_top_",colnames(data)))]
data <- data[,-(grep("_roots_",colnames(data)))]
```

Keep only unique values:

```{r UniqueValuesProvenancesP1}
data <- unique(data)
```


## Exploratory analyses


> Checking NAs

```{r noNAsProvenancesP1}
sapply(data, function(x) sum(is.na(x)))
```


> Mean-centered variables

```{r ScalingVariablesProvenancesP1}
data[,paste0(grep("_prov",names(data),value = T),".sc")] <- scale(data[,grep("_prov",names(data),value = T)])
```


```{r SquareCorrplotDataProvenancesP1, fig.height=12}
cor <- cor(data[,grep("_prov.sc",names(data),value = T)])
corrplot::corrplot(cor,method="number")
```


```{r CorrplotTriangleProvenancesP1, fig.height=12}
# matrix of the p-value of the correlation
p.mat <- corpmat(cor)

corrplot::corrplot(cor, method="color", col=col(200),  
                   type="upper", order="hclust", 
                   addCoef.col = "black", # Add coefficient of correlation
                   tl.col="black", tl.srt=45, #Text label color and rotation
                   # Combine with significance
                   p.mat = p.mat, sig.level = 0.01, insig = "blank", 
                   # hide correlation coefficient on the principal diagonal
                   diag=FALSE)
```



```{r PCAProvenancesP1, fig.height=12}
pca <- prcomp(data[,grep("_prov.sc",names(data),value = T)], center = TRUE,scale. = TRUE)
ggbiplot(pca,varname.size =4) +  ylim(-4.5, 2.5) +    xlim(-3, 3) +  
   theme_minimal(base_size = 25) +
  theme(plot.title = element_text(size=18))
```

## Selected variables

We didn't keep bio13 because it was highly correlated to bio12. And we didn't keep `bio6` because it was correlated to `bio1`.

```{r CorrBio12Bio13Bio5Bio6ProvenancesP1}
cat(" data:","\n","corr bio12-bio13: ",cor(data$bio12_prov.sc,data$bio13_prov.sc), "\n",
    "corr bio5-bio6: ",cor(data$bio1_prov.sc,data$bio6_prov.sc),"\n","\n")
```

Selected variables:

- `bio1_prov.s` (*mean.temp* in the manuscript): the average of the annual daily mean temperature (°C).

- `bio5_prov.sc` (*max.temp* in the manuscript): the average of the maximum temperature of the warmest month (°C).

- `bio12_prov.sc` (*min.pre* in the manuscript): the average of the precipitation of the driest month (mm).

- `bio14_prov.sc` (*mean.pre* in the manuscript): the average of the annual precipitation (mm).


```{r SelectedVariablesProvenancesP1}
select.var <- paste0("bio",c(1,5,12,14),"_prov.sc")
```

```{r CorrelationsProvenancesP1}
cor(data[,select.var])
```

```{r VIFProvenancesP1}
vif(data[,select.var]) 
```

## Matrices of environmental similarity

2 options:

- A matrix of Euclidean distance following @thomsonGuideUsingMultiplematrix2018.
    
- A variance-covariance matrix following @jarquinReactionNormModel2014.


```{r PrepareDatasetProvenancesP1}
dat <- unique(data[,c("prov",select.var)])
row.names(dat) <- dat$prov
dat$prov <- NULL 
round(dat,3)
```

### Matrix of Euclidean distance 

Following @thomsonGuideUsingMultiplematrix2018.

I did not select this option in the manuscript as I did not know how to include a euclidean matrix in `brms`. 

```{r EucliMatrixProvenancesP1}
euclimat <-as.matrix(dist(dat, method="euclidean", diag=TRUE,upper=TRUE))
round(euclimat,3)[1:10,1:10]

# then scales so that the values are between 0 and 1
euclimat <- 1- euclimat/max(euclimat, na.rm=TRUE)
round(euclimat,3)[1:10,1:10]
```

```{r SaveEucliMatrixProvenancesP1}
# saveRDS(euclimat, file = "../../data/EnvMat/EucliMatrixProvenancesP1.rds")
```


### Variance-covariance matrix

Following @jarquinReactionNormModel2014.

This is the option used in the mannuscript.

```{r VarCovMatrixProvenancesP1,fig.height=13,fig.width=13}
varmat <-as.matrix(var(t(dat)))
round(varmat,3)[1:10,1:10]
heatmap(varmat)

heatmap.2(as.matrix(varmat),scale="none",
          col=brewer.pal(11,"RdBu"),
          trace="none",key=F,
          cexRow=1.5,cexCol=1.5,
          margins=c(12,8))
```

```{r SaveHeatmapVarCovMatrixProvenancesP1, echo=F,message=F}
png("../../figs/SuppInfo/ExploringData/VarCovMatrixProvenancesP1.png",width = 1200,height=1200)
heatmap.2(as.matrix(varmat),scale="none",
          col=brewer.pal(11,"RdBu"),
          trace="none",key=F,
          cexRow=1.5,cexCol=1.5,
          margins=c(12,8))
dev.off() 
```

```{r PosDefVarCovMatrixProvenancesP1}
is.symmetric.matrix(varmat)
is.positive.definite(varmat)
varmat.pd <- nearPD(varmat)$mat
all.equal(as.matrix(varmat.pd),varmat)
```


```{r SavePosDefVarCovMatrixProvenancesP1}
# saveRDS(varmat.pd, file = "../../data/cleaned-data/EnvMat/VarCovMatProvenancesP1.rds")
```




# Climatic similarity between provenances (P2)

## Load and clean data


```{r LoadDataProvenancesP2}
data <- readRDS(file="../../data/TrainP1.RDS")
```


> Keeping the columns of interest

```{r KeepColumnOfInterestProvenancesP2}
data <- data[,c("prov",grep("_prov",colnames(data),value=T))]
data <- data[,-(grep("tude",colnames(data)))]

# Removing soil variables
data <- data[,-(grep("_sub_",colnames(data)))]
data <- data[,-(grep("_top_",colnames(data)))]
data <- data[,-(grep("_roots_",colnames(data)))]
```

Keep only unique values:

```{r UniqueValuesProvenancesP2}
data <- unique(data)
```


## Exploratory analyses


> Checking NAs

```{r noNAsProvenancesP2}
sapply(data, function(x) sum(is.na(x)))
```


> Mean-centered variables

```{r ScalingVariablesProvenancesP2}
data[,paste0(grep("_prov",names(data),value = T),".sc")] <- scale(data[,grep("_prov",names(data),value = T)])
```


```{r SquareCorrplotDataProvenancesP2, fig.height=12}
cor <- cor(data[,grep("_prov.sc",names(data),value = T)])
corrplot::corrplot(cor,method="number")
```


```{r CorrplotTriangleProvenancesP2, fig.height=12}
# matrix of the p-value of the correlation
p.mat <- corpmat(cor)

corrplot::corrplot(cor, method="color", col=col(200),  
                   type="upper", order="hclust", 
                   addCoef.col = "black", # Add coefficient of correlation
                   tl.col="black", tl.srt=45, #Text label color and rotation
                   # Combine with significance
                   p.mat = p.mat, sig.level = 0.01, insig = "blank", 
                   # hide correlation coefficient on the principal diagonal
                   diag=FALSE)
```



```{r PCAProvenancesP2, fig.height=12}
pca <- prcomp(data[,grep("_prov.sc",names(data),value = T)], center = TRUE,scale. = TRUE)
ggbiplot(pca,varname.size =4) +  ylim(-4.5, 2.5) +    xlim(-3, 3) +  
   theme_minimal(base_size = 25) +
  theme(plot.title = element_text(size=18))
```

## Selected variables

We didn't keep bio13 because it was highly correlated to bio12. And we didn't keep `bio6` because it was correlated to `bio1`.

```{r CorrBio12Bio13Bio5Bio6ProvenancesP2}
cat(" data:","\n","corr bio12-bio13: ",cor(data$bio12_prov.sc,data$bio13_prov.sc), "\n",
    "corr bio5-bio6: ",cor(data$bio1_prov.sc,data$bio6_prov.sc),"\n","\n")
```

Selected variables:

- `bio1_prov.s` (*mean.temp* in the manuscript): the average of the annual daily mean temperature (°C).

- `bio5_prov.sc` (*max.temp* in the manuscript): the average of the maximum temperature of the warmest month (°C).

- `bio12_prov.sc` (*min.pre* in the manuscript): the average of the precipitation of the driest month (mm).

- `bio14_prov.sc` (*mean.pre* in the manuscript): the average of the annual precipitation (mm).


```{r SelectedVariablesProvenancesP2}
select.var <- paste0("bio",c(1,5,12,14),"_prov.sc")
```

```{r CorrelationsProvenancesP2}
cor(data[,select.var])
```

```{r VIFProvenancesP2}
vif(data[,select.var])
```

## Matrices of environmental similarity

2 options:

- A matrix of Euclidean distance following @thomsonGuideUsingMultiplematrix2018.
    
- A variance-covariance matrix following @jarquinReactionNormModel2014.


```{r PrepareDatasetProvenancesP2}
dat <- unique(data[,c("prov",select.var)])
row.names(dat) <- dat$prov
dat$prov <- NULL 
round(dat,3)
```

### Matrix of Euclidean distance 

Following @thomsonGuideUsingMultiplematrix2018.

I did not select this option in the manuscript as I did not know how to include a euclidean matrix in `brms`. 

```{r EucliMatrixProvenancesP2}
euclimat <-as.matrix(dist(dat, method="euclidean", diag=TRUE,upper=TRUE))
round(euclimat,3)[1:10,1:10]

# then scales so that the values are between 0 and 1
euclimat <- 1- euclimat/max(euclimat, na.rm=TRUE)
round(euclimat,3)[1:10,1:10]
```

```{r SaveEucliMatrixProvenancesP2}
# saveRDS(euclimat, file = "../../data/EnvMat/EucliMatrixProvenancesP2.rds")
```


### Variance-covariance matrix

Following @jarquinReactionNormModel2014.

This is the option used in the mannuscript.

```{r VarCovMatrixProvenancesP2,fig.height=13,fig.width=13}
varmat <-as.matrix(var(t(dat)))
round(varmat,3)[1:10,1:10]
heatmap(varmat)

heatmap.2(as.matrix(varmat),scale="none",
          col=brewer.pal(11,"RdBu"),
          trace="none",key=F,
          cexRow=1.5,cexCol=1.5,
          margins=c(12,8))
```

```{r SaveHeatmapVarCovMatrixProvenancesP2, echo=F,message=F}
png("../../figs/SuppInfo/ExploringData/VarCovMatrixProvenancesP2.png",width = 1200,height=1200)
heatmap.2(as.matrix(varmat),scale="none",
          col=brewer.pal(11,"RdBu"),
          trace="none",key=F,
          cexRow=1.5,cexCol=1.5,
          margins=c(12,8))
dev.off() 
```

```{r PosDefVarCovMatrixProvenancesP2}
is.symmetric.matrix(varmat)
is.positive.definite(varmat)
varmat.pd <- nearPD(varmat)$mat
all.equal(as.matrix(varmat.pd),varmat)
```


```{r SavePosDefVarCovMatrixProvenancesP2}
# saveRDS(varmat.pd, file = "../../data/cleaned-data/EnvMat/VarCovMatProvenancesP2.rds")
```

`r knitr::opts_chunk$set(eval = F)`

# References